### 2020-11-11 C.analysis/markdowns/Experiment_1.Rmd
Add the drake workflow scripts

---
title: "Experiment 1"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: true
    toc_float: true
---

# Description

In experiment 1 we apply the different FoRTE girdling treatments to the ED forest stand. This stand starts from near bare ground starting in 1900. After a series of tests, we had to exclude pines from the ED configuration.

See [GitHub issue](https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77#issue-687269729) for some more details.

Observations

* NPP winter pattern is interesting
* By 2030 NPP hasn't recovered 
* There is a small pattern in the Rh

Questions

* Based on discussion with Lisa do we want to take a look at aboveground bark biomass? Would that be something easy to compare with Max's dendrology band measurements? 
* Would it make sense to report values as difference harvest 45% - harvest 0%? Or the difference from a refernce value for a single year? 
* Should we use ED output to calculate resillence? 
* Should we run ED with climatological inputs?

### Set Up

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# The defaults for the chunks
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 3)

library(ggplot2)
library(magrittr)
library(data.table)
library(cowplot)
library(lubridate)
library(purrr)

BASE_DIR <- here::here()
THEME    <- theme_bw(base_size = 15)
```

Import the results

```{r}
# Note that this is the object created with the drake frame work, there should be some better way to trigger this. 
exp_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', "exp-1.rds"))
exp_object <- map(exp_object, function(x){ x[year < 2030, ] }) # Note the year 2030 is not a complete run 
```

Event information

```
<year> 2019 </year>
<doy> 141 </doy>
```

# Baseline Scenario

```{r, fig.width = 3, fig.height = 3}
ggplot() + 
  geom_point(data = exp_object$NPP[scn == 'harvest_0', ], aes(year, value)) +
  THEME +
  labs(title ='NPP',
       x = 'Year', 
       y = unique(exp_object$NPP[scn == 'harvest_0', ][['unit']])) -> 
  NPP_plot

ggplot() + 
  geom_point(data = exp_object$LAI[scn == 'harvest_0', ], aes(datetime, value)) + 
  THEME + 
  labs(title ='LAI',
       x = 'Year', 
       y = unique(exp_object$LAI[scn == 'harvest_0', ][['unit']])) -> 
  LAI_plot

ggplot() + 
  geom_point(data = exp_object$NEP[scn == 'harvest_0', ], aes(year, value)) +
  THEME + 
  labs(title ='NEP',
       x = 'Year', 
       y = unique(exp_object$NEP[scn == 'harvest_0', ][['unit']])) -> 
  NEP_plot

ggplot() + 
  geom_point(data = exp_object$ABG[scn == 'harvest_0', ], aes(year, value)) + 
  THEME + 
  labs(title ='ABG',
       x = 'Year', 
       y = unique(exp_object$ABG[scn == 'harvest_0', ][['unit']])) -> 
  ABG_plot

plot_grid(ABG_plot, NPP_plot, NEP_plot, LAI_plot)

```

The base line scenario looks good to me it is intresting just how much smoother the output looks when we swtich to the climatological record instead of weather, do we want to only use the climatological values.

# Experiment 1

## Above Ground Biomass

So the differences in total above ground biomass between the different sencaiors is faily small and is really only visible when we zoom in on the disturbance and recovery. What is intresting is that the as we would expect the no harvest scenario has a larger biomass until September and then the AGB starts to even out, is pretty similar during winter but then with the next growing season the patches with the more extreme treatments have a larger biomass.

```{r}
exp_object$ABG$datetime <- date(exp_object$ABG$datetime)
data <- exp_object$ABG[  , value := sum(value), by = c("datetime", "scn", "unit", "year")]

data %>%  
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_point(aes(datetime, value, color = scn, group = interaction(scn, pft_name))) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
```

```{r}
data[year %in% 2018:2020 & scn != 'harvest_0', ] %>% 
  .[datetime >= date('2019-05-01')] %>%  
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn, group = interaction(scn, pft_name)), size = 1) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
```

## NPP

Annual values, what is odd is that the harvest 0% has NPP is more negative than the other treatments? Is that because of reading in the 0 harvest? Also does the monthly NPP have any other patterns?

```{r}
exp_object$NPP %>% 
  .[year %in% 2018:20205, ] %>% 
  ggplot() + 
  geom_vline(aes(xintercept = 2019), color ='black') + 
  geom_line(aes(year, value, color = scn, group = interaction(scn, pft_name))) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NPP$unit), title = 'Annual NPP')
```

So the annual is some what hard to see the the pattern with it but after a decade there the NPP has not returned to the annual values.

```{r}
h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))

h0 <- as.data.table(h0_object$df_cohort[[1]])[  ,list(datetime, MMEAN_NPP_CO, PFT, DBH, NPLANT)]
h45 <- as.data.table(h45_object$df_cohort[[1]])[  ,list(datetime, MMEAN_NPP_CO, PFT, DBH, NPLANT)]
h65 <- as.data.table(h65_object$df_cohort[[1]])[  ,list(datetime, MMEAN_NPP_CO, PFT, DBH, NPLANT)]
h85 <- as.data.table(h85_object$df_cohort[[1]])[  ,list(datetime, MMEAN_NPP_CO, PFT, DBH, NPLANT)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

NPP <- rbind(h0, h45, h65, h85)
# Convert the NPP units from per plant to unit area
NPP$value <- NPP$MMEAN_NPP_CO * NPP$NPLANT
NPP$unit  <- unique(exp_object$NPP$unit)
NPP <- NPP[, c("datetime", "PFT", "scn", "value", 'unit')]
NPP$year <- lubridate::year(NPP$datetime)
NPP$datetime <- date(NPP$datetime)
NPP_monthly <- NPP 
```

```{r}
NPP_monthly <- NPP_monthly[, value := sum(value), by = c("datetime", "scn", "unit", "year")]

NPP_monthly %>% 
  ggplot(aes(datetime, value, color = scn)) + 
  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
```

```{r}
NPP_monthly %>% 
  .[year %in% 2018:2030] %>%  
  ggplot(aes(datetime, value, color = scn)) + 
  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
```

Would it make sense to look at the differrence between the baseline and treatment scenarios?

Observations

* Even after 12 years the stand has not recovered in term of NPP values 
* Also check out the difference in the winter NPP! I wasn't expecting that

## NEP

```{r}
exp_object$NEP %>% 
  dplyr::filter(year >= 2017 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = 2019, color = 'grey', size = 2) + 
  geom_line(aes(year, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Annual NEP')
```

```{r}
h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))

h0 <- as.data.table(h0_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h45 <- as.data.table(h45_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h65 <- as.data.table(h65_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
h85 <- as.data.table(h85_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

NEP <- rbind(h0, h45, h65, h85)
NEP$unit  <- unique(exp_object$NEP$unit)
NEP <- NEP[, c("datetime", "scn", "value" = "MMEAN_NEP_PY")]
names(NEP) <- c("datetime", "scn", "value")
NEP$year <- lubridate::year(NEP$datetime)
NEP$datetime <- date(NEP$datetime)
```

```{r}
NEP %>% 
  dplyr::filter(year >= 2019 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Monthly NEP')
```

## LAI

```{r}
exp_object$LAI$year <- year(exp_object$LAI$datetime)
exp_object$LAI %>%
  dplyr::filter(year >= 2019 & year <= 2021) %>% 
  ggplot() + 
  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
  geom_line(aes(datetime, value, color = scn)) + 
  THEME + 
  labs(x = 'Year', y = unique(exp_object$LAI$unit), title = 'Monthly LAI') + 
  facet_wrap('pft_name', ncol = 1)
```

## Soil Respiration

```{r}
h0 <- as.data.table(h0_object$df_scalar[[1]])[  ,list(datetime, MMEAN_RH_PY)]
h45 <- as.data.table(h45_object$df_scalar[[1]])[  ,list(datetime, MMEAN_RH_PY)]
h65 <- as.data.table(h65_object$df_scalar[[1]])[  ,list(datetime, MMEAN_RH_PY)]
h85 <- as.data.table(h85_object$df_scalar[[1]])[  ,list(datetime, MMEAN_RH_PY)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

RH    <- rbind(h0, h45, h65, h85)
units <- 'kg/m2/yr'

RH %>% 
ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time')
```

```{r}
RH[ , year := year(datetime)] %>% 
  .[year %in% 2018:2020, ] %>% 
  ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Heterotrophic Respiration', 
       y = units,
       x = 'Date Time')
```

## Mortality

Honestly, I don't really know what the morality output is so might not be easy to interpret but it looks like by 2020 the morality related to the disturbance events has ended.

```{r}
h0 <- as.data.table(h0_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h45 <- as.data.table(h45_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h65 <- as.data.table(h65_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
h85 <- as.data.table(h85_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]

h0$scn  <- 'harvest_0'
h45$scn <- 'harvest_45'
h65$scn <- 'harvest_65'
h85$scn <- 'harvest_85'

mortality    <- rbind(h0, h45, h65, h85)
units <- '1/yr'

mortality[ ,year := year(datetime)] %>% 
  .[year %in% 2018:2021, ] %>% 
ggplot(aes(datetime, mmean_mort_rate_co_carbon, color = scn)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Monthly Mean Mortlaity Rate per Cohort', 
       y = units,
       x = 'Date Time')
```

<br>

****

```{r}
sessionInfo()
```

### 2020-11-11 C.analysis/markdowns/Inputs_TestRun.Rmd
Add the drake workflow scripts

---
title: 'FoRTE-Disturbance: Inputs & Test Run'
author: "Kalyn Dorheim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Objective & Background

This document serves to document the FoRTE ED disturbance baseline run. It offers us an opportunity to take a look at the meteorological inputs. Things we need to think about while looking at the results

* Does the meteorological input look good? 
* Do we like how the output is reported? Units? Do we want to look at the mid summer value, annual totals? 
* Are there any concerns we may have? What like the negative NEP?

\
\

Met Data List Name

> 
* dlwrf — downward long wave radiation (W m-2)
* nbdsf — near infrared beam downward solar radiation (W m-2)
* nddsf — near IR diffuse downward solar radiation (W m-2)
* vbdsf — visible beam downward solar radiation (W m-2)
* vddsf — visible diffuse downward solar radiation (W m-2)
* prate — precipitation rate (kgH2O m-2 s-1)
* pres — atmospheric pressure (Pa)
* hgt — geopotential height (m)
* ugrd — zonal wind (m s-1)
* vgrd — meridional wind (m s-1)
* sh — specific humidity (kgH2O kgAir-1)
* tmp — air temperature (K)

# Set Up

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# The defaults for the chunks
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)

# Load required libraries
library(magrittr)
library(ggplot2)
library(data.table)
library(lubridate)
library(ed4forte)
library(cowplot)
library(fortedata)

# Set up the dirs 
BASE_DIR <- here::here()
INPUT <- file.path(BASE_DIR, "A.inputs")
OUTPUT_DIR <- file.path(BASE_DIR, "ED-outputs")
THEME <- theme_bw()
```

```{r}
data <- as.data.table(read.csv(file.path(INPUT, 'monthly_mean_met.csv'), stringsAsFactors = FALSE))
data$date <- ymd(paste(data$year, data$month, '01', sep = '/'))
```

```{r, warning=FALSE}
long_met_data <- melt(data, measure.vars = c("nbdsf", "nddsf", "vbdsf", "vddsf",
                                             "prate", "dlwrf", "pres", "hgt", "ugrd", 
                                             "vgrd", "sh", "tmp"),
                      variable.name = 'variable', value.name = 'value')
long_met_data <- long_met_data[, list(date, variable, value, year)]

var_info <- data.table(variable = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K'))
long_met_data <- long_met_data[var_info, on = 'variable']
long_met_data <- long_met_data[variable != 'hgt', ] # exclude the geo spatial height on cause it doesn't 
```

# Inputs

```{r}
long_met_data %>% 
  ggplot() + 
  geom_line(aes(date, value)) + 
  facet_wrap("variable", scales = 'free') + 
  labs(title = 'Monthly Time Series of ED Met Data Inputs') + 
  geom_vline(xintercept = date('1979-01-01'), color = 'red') + 
  geom_vline(xintercept = date('2015-01-01'), color = 'blue') + 
  THEME
```

The red line indicate when the generated data historical Climate data ends. The blue line indicates when the dermatological record is used starting in 2015.

The monthly data is rather noisy and it can be difficult to look at, calculate the annual mean.

```{r}
annual_data <- long_met_data[ , list( 'value' = mean(value)), by = list(year, variable)]
```

```{r}
ggplot(annual_data) + 
  geom_line(aes(year, value)) + 
  facet_wrap("variable", scales = 'free') + 
  labs(title = 'Annual ED Met Data Inputs') +  
  geom_vline(xintercept = 1979, color = 'red') + 
  geom_vline(xintercept = 2015, color = 'blue') + 
  THEME
```

# Test Run Results

As part of the development of the ED FoRTE experiment runs I did a complete default run using all the met inputs. The outputs we are interested in are monthly NPP, NEE, LAI, biomass by cohort.

```{r}
# Read in the ED data, it is in a very particular format, in order to plot there will need to be some 
# data manipulation. 
ed_data <- readRDS(file.path(OUTPUT_DIR, "test", "baseline.rds"))
PFTs <- c(6, 8, 9, 10, 11, 7) # The PFTS to keep, some of the ED outputs report all of the PFTs.

pft_names <- data.table(pft = PFTs, 
                        pft_name = c('Northern North American pines', 
                                     'Late-successional North American conifers', 
                                     'Temperate broadleaf, early successional',  
                                     'Temperate broadleaf, mid-successional', 
                                     'Temperate broadleaf, late successional', 
                                     'Southern North American pines'))

```

monthly NPP, NEE, LAI, biomass by cohort

### Biomass

In these plots I've only included the values from June to simplify the plots. Otherwise it is near impossible to differentiate between the pfts.

```{r}
out <- as.data.table(ed_data$df_cohort[[1]])[  ,list(datetime, AGB_CO, PFT, DBH, NPLANT)]

# Add a unique identifier for each of the cohorts (reminder that cohorts = unique species at DBH)
split(out, out$datetime) %>%  
  lapply(function(x){

    x$CO <- LETTERS[1:nrow(x)]
    return(x)

  }) %>% 
  rbindlist() -> 
  out

out <- out[pft_names, on = c('PFT' = 'pft')]
out$year <- year(out$datetime)
out$month <- month(out$datetime)
out$variable <- "AGB_CO"
out <- out[as.data.table(ed2_variable_info())[, list(variable, description, unit)], on = 'variable',  nomatch=0]
out$value <- out$NPLANT * out$AGB_CO
out$unit <- gsub(pattern = '/plant', replacement = '/m2', out$unit)

out[month == 8]  %>% 
  ggplot(aes(year, value, color = DBH, shape = pft_name)) + 
  geom_point() + 
  THEME + 
  labs(title = 'Mid Summer Above Ground Biomass by Cohort (pft x DBH)',
       subtitle = 'Complete Run',
       y = unique(out$unit)) + 
  geom_vline(xintercept = 1979, color = 'red') + 
  geom_vline(xintercept = 2015, color = 'blue') ->
  full_run

out[year %in% 1900:1920 & month == 8] %>% 
  ggplot(aes(datetime, AGB_CO, color = DBH, shape = pft_name)) + 
  geom_point() + 
  THEME + 
  labs(title = 'Mid Summer Above Ground Biomass by Cohort (pft x DBH)',
       subtitle = 'Early Years',
       y = unique(out$unit)) + 
  theme(legend.position = 'none') ->
  early_years

plot_grid(early_years, full_run, rel_widths = c(0.3,  0.7))
```

\
\

### NPP

```{r}
out <- as.data.table(ed_data$df_cohort[[1]])[  ,list(datetime, MMEAN_NPP_CO, PFT, DBH, NPLANT)]
out <- out[pft_names, on = c('PFT' = 'pft')]
out$year <- year(out$datetime)
out$month <- month(out$datetime)
out$variable <- 'MMEAN_NPP_CO'
out <- out[as.data.table(ed2_variable_info())[, list(variable, description, unit)], on = 'variable',  nomatch=0]
```

```{r}
out$value <- out$MMEAN_NPP_CO * out$NPLANT
out$unit <-gsub(pattern = '/pl', replacement = '/m2', out$unit)

npp_to_plot <-out
out <- out[, list(value = mean(MMEAN_NPP_CO, na.rm = TRUE)), by = list(pft_name, year, variable, unit, description)]
```

```{r}
out %>% 
  ggplot(aes(year, value, color = pft_name)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Average Annual NPP',
       subtitle = 'PFT',
       y = unique(out$unit)) + 
  geom_vline(xintercept = 1979, color = 'red') + 
  geom_vline(xintercept = 2015, color = 'blue') 
```

\
\

### NEP

Because NEP looks at the ecosystem is is not reported on the cohort level.

```{r}
out <- as.data.table(ed_data$df_scalar[[1]])[  ,list(datetime, MMEAN_NEP_PY)]
out$datetime <- ymd(out$datetime)
out$year <- year(out$datetime)
out$month <- month(out$datetime)
out$variable <- 'MMEAN_NEP_PY'
out <- out[as.data.table(ed2_variable_info())[, list(variable, description, unit)], on = 'variable',  nomatch=0]

out <- out[, list(value = mean(MMEAN_NEP_PY, na.rm = TRUE)), by = list(year, variable, unit, description)]

```

```{r}
out %>%  
  ggplot(aes(year, value)) + 
  geom_line() + 
  THEME + 
  labs(title = 'Annual NEP',
       subtitle = 'Stand NEP',
       y = unique(out$unit)) + 
  theme(legend.position = 'none') + 
  geom_vline(xintercept = 1979, color = 'red') + 
  geom_vline(xintercept = 2015, color = 'blue') 
```

\
\

### LAI

```{r}
out <- as.data.table(ed_data$df_pft[[1]])[  ,list(datetime, MMEAN_LAI_PY, pft)]
out <- dplyr::filter(out, pft %in% c(10, 11, 9))
out$datetime <- ymd(out$datetime)
out$year     <- year(out$datetime)
out$month    <- month(out$datetime)
out$variable <- 'MMEAN_LAI_CO'
out <- out[as.data.table(ed2_variable_info())[, list(variable, description, unit)], 
           on = 'variable',  nomatch=0]
out <- out[pft_names, on = c('pft')]
out <- dplyr::filter(out, pft %in% c(10, 11, 9))

```

```{r}
out %>% 
 # dplyr::filter(year < 1905) %>%  
  ggplot(aes(ymd(datetime), MMEAN_LAI_PY, color = pft_name)) + 
  geom_line() + 
  THEME +
  labs(title = 'Leaf area index',
       subtitle = 'PFT, no pines',
       y = unique(out$unit), 
       x = 'Mid Summer Values') + 
  # theme(legend.position = 'none') +
  geom_vline(xintercept = ymd('1979-01-01'), color = 'red') + 
  geom_vline(xintercept = ymd('2015-01-01'), color = 'blue') + 
 # coord_cartesian(ylim = c(0, 0.05)) + 
  NULL
```

\
\
\

****

```{r}
sessionInfo()
```

### 2020-11-11 C.analysis/functions.R

@@ -1,0 +1,245 @@
+# Custom functions for the drake workflow that is used to process the 
+# monthly ED output files into data frames that are easy to work with. 
+
+
+# Add the ED pft names of the data table 
+#   Args  
+#     dt: the data table of ED information, it must have PFT column 
+#   Return: a data table with the pft names 
+add_pft_names <- function(dt){
+  
+  # Check the inputs 
+  assert_that(is.data.table(dt))
+  assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
+  
+  # Data table of the plant functional types in ED. 
+  pft_names <- data.table(pft = c(6, 8, 9, 10, 11, 7) , 
+                          pft_name = c('Northern North American pines', 
+                                       'Late-successional North American conifers', 
+                                       'Temperate broadleaf, early successional',  
+                                       'Temperate broadleaf, mid-successional', 
+                                       'Temperate broadleaf, late successional', 
+                                       'Southern North American pines'))
+  
+  
+  # Add the pft names to the pft data table. 
+  named_dt <- na.omit(dt[pft_names, on = c("PFT" = "pft")])
+  return(named_dt)
+  
+}
+
+# KALYN DOING THIS ONE WRONG, IT IS SUPPOSED TO BE THE OTHER WAY 
+# Add the UMBS species and the pft name to a ED data table
+#   Args 
+#     dt: a data table of ED output that contains the PFT column
+# Return: a data table of ED output with the PFT names and UMBS species 
+# UMBS_to_ED <- function(dt){
+#   
+#   # Check the inputs 
+#   assert_that(is.data.table(dt))
+#   assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
+#   
+#   # A Data table that maps the ED PFT to the UMBS species 
+#   mid_success       <- data.table(pft = 10, UMBS = c('ACRU','ACPE',  'FAGR', 'ACSA'), stringsAsFactors = FALSE)
+#   temp_broadl_late  <- data.table(pft = 11, UMBS = c('QURU', 'TCSA'), stringsAsFactors = FALSE)
[…]

### 2020-11-11 C.analysis/packages.R

@@ -1,0 +1,7 @@
+# Load all your packages before calling make().
+
+library(drake)
+library(tibble)
+library(data.table)
+library(lubridate)
+library(assertthat)

### 2020-11-11 C.analysis/plan.R

@@ -1,0 +1,153 @@
+# Define drake plans that are used to process/format the ED experiment results. 
+
+# 0. Set Up ------------------------------------------------------------------
+source(here::here('C.analysis', 'packages.R'))
+source(here::here('C.analysis', 'functions.R'))
+
+ED_OUTPUT_DIR <- here::here('ED-outputs')
+
+# 1. Test Results --------------------------------------
+results_test <- drake_plan(
+
+  # Baseline
+  test_baseline_ed_object = target(readRDS(file.path(ED_OUTPUT_DIR, "test", "baseline.rds"))), 
+  test_baseline_data = target({
+    scn_name <- 'test-baseline'
+    data <- list(NPP = process_NPP(test_baseline_ed_object, scn_name), 
+                 LAI = process_LAI(test_baseline_ed_object, scn_name), 
+                 NEP = process_NEP(test_baseline_ed_object, scn_name), 
+                 ABG = process_ABG(test_baseline_ed_object, scn_name))
+    data
+  }),
+  
+  # Save Output 
+  target({
+    out <- list(NPP = ed_format_list('NPP', list(test_baseline_data)),
+                LAI = ed_format_list('LAI', list(test_baseline_data)),
+                NEP = ed_format_list('NEP', list(test_baseline_data)),
+                ABG = ed_format_list('ABG', list(test_baseline_data)))
+    saveRDS(object = out, file = file.path(ED_OUTPUT_DIR, "test-data.rds"))
+    
+  })
+)
+
+# 2. Exp 1 ------------------------------------------------------------------------------
+results_exp1 <- drake_plan(
+  
+  # Baseline 
+  harvest_0_object =  target(readRDS(file.path(ED_OUTPUT_DIR, "test", "baseline.rds"))), 
+  harvest_0_data = target({
+    scn_name <- 'harvest_0'
+    data <- list(NPP = process_NPP(harvest_0_object, scn_name), 
+                 LAI = process_LAI(harvest_0_object, scn_name), 
+                 NEP = process_NEP(harvest_0_object, scn_name), 
+                 ABG = process_ABG(harvest_0_object, scn_name))
+    data
+  }), 
+  
+  # 45 % FoRTE Treatment 
+  harvest_45_object = target(readRDS(file.path(ED_OUTPUT_DIR, "exp-1", "harvest_45.rds"))), 
+  harvest_45_data = target({
[…]

### 2021-01-14 A.inputs/3.create_eventfiles.R
Running experiment one

@@ -1,0 +1,216 @@
+# 0. Set Up ---------------------------------------------------------------------------
+library(xml2)
+library(assertthat)
+library(lubridate)
+library(data.table)
+
+# Define the paths 
+BASE_DIR <- here::here()
+WRITE_TO <- file.path(BASE_DIR, 'A.inputs', 'events')
+
+
+# 1. Define functions --------------------------------------------------------------------------------
+# Modify an ED harvest xml file 
+# Args
+#   table: a data frame of the year, doy, agb_frac, bgb_frac, fol_frac, and stor_frac to add to the harvest xml 
+#   xml_name: the path to write the xml file out to 
+#   xml_template_file: default set to the template in the A.inputs/events file. This is a 0 disturabnce event. 
+# Return: the name of the xml file that was created. 
+modify_xml <- function(table, xml_name, xml_template_file = here::here('A.inputs', 'events', 'harvest_template.xml')){
+  
+  # Check the function arguments
+  assert_that(has_name(table, c("year", "doy", "agb_frac", "bgb_frac", "fol_frac", "stor_frac")))
+  assert_that(sum(is.na(table)) == 0)
+  assert_that(all(is.double(unlist(table))))
+  assert_that(dir.exists(dirname(xml_name)))
+  assert_that(file.exists(xml_template_file))
+  assert_that(grepl(pattern = '.xml', x = xml_name))
+  
+  # Import the xml file 
+  xml_template <- read_xml(xml_template_file)
+  
+  # Extract the childe node, this is the harvest event template that will be used to 
+  # expand the template xml so there is a harvest entry for each entry in the table. 
+  child_node <- xml_child(xml_template)
+  
+  # Add sibling nodes to the xml template for all of then entries in the harvest file. 
+  # Because the xml file already contains a harvest event node add 1 less node than 
+  # listed in the the table. 
+  for(index in 1:(nrow(table) - 1)){
+    xml_add_sibling(child_node, child_node, where = "after")
+  }
+  
[…]

### 2021-01-14 A.inputs/3.modify_create_eventfiles.R

@@ -1,97 +1,0 @@
-# 0. Set Up ---------------------------------------------------------------------------
-library(xml2)
-library(assertthat)
-library(lubridate)
-
-# Define the paths 
-BASE_DIR <- here::here()
-WRITE_TO <- file.path(BASE_DIR, 'A.inputs', 'events')
-
-
-# 1. Define functions --------------------------------------------------------------------------------
-# Modify an ED harvest xml file 
-# Args
-#   table: a data frame of the year, doy, agb_frac, bgb_frac, fol_frac, and stor_frac to add to the harvest xml 
-#   xml_name: the path to write the xml file out to 
-#   xml_template_file: default set to the template in the A.inputs/events file. This is a 0 disturabnce event. 
-# Return: the name of the xml file that was created. 
-modify_xml <- function(table, xml_name, xml_template_file = here::here('A.inputs', 'events', 'harvest_template.xml')){
-  
-  # Check the function arguments
-  assert_that(has_name(table, c("year", "doy", "agb_frac", "bgb_frac", "fol_frac", "stor_frac")))
-  assert_that(sum(is.na(table)) == 0)
-  assert_that(all(is.double(unlist(table))))
-  assert_that(dir.exists(dirname(xml_name)))
-  assert_that(file.exists(xml_template_file))
-  assert_that(grepl(pattern = '.xml', x = xml_name))
-  
-  # Import the xml file 
-  xml_template <- read_xml(template_xml_file)
-  
-  # Extract the childe node, this is the harvest event template that will be used to 
-  # expand the template xml so there is a harvest entry for each entry in the table. 
-  child_node <- xml_child(xml_template)
-  
-  # Add sibling nodes to the xml template for all of then entries in the harvest file. 
-  # Because the xml file already contains a harvest event node add 1 less node than 
-  # listed in the the table. 
-  for(index in 1:(nrow(table) - 1)){
-    xml_add_sibling(child_node, child_node, where = "after")
-  }
-  
-  # Now that the xml has been modified to the appropriate size convert to a list. 
-  xml_list <- as_list(xml_template)
-  
-  # Now modify 
-  for(i in 1:nrow(table)){
-    
[…]

### 2021-01-14 B.ed/1.Experiment-1.R

@@ -22,8 +22,10 @@ source(to_source)
 # Start by defining the ED run information. 
 IYEARA <- 1900
 IYEARZ <- 2030
-cases  <- c('harvest_0',  'harvest_45','harvest_65', 'harvest_85')
-event_file <- file.path(EVENT_DIR, paste0(cases, '.xml')) # There should be an event file for each of FoRTE treatment groups.
+
+# Find the event files
+event_file <- list.files(EVENT_DIR, pattern = '1day_above.xml', full.names = TRUE)
+cases      <- gsub(pattern = '.xml', replacement = '', x = basename(event_file))
 assertthat::assert_that(all(file.exists(event_file)))
 # Important! When reading in the event_file paths they must be written out as a string, 
 # otherwise ED will fail but will throw an uninformative error message realted to 

### 2021-01-14 B.ed/2.post_processing.R
cleaning up

@@ -14,24 +14,22 @@ stopifnot(file.exists(to_source))
 source(to_source)
 
 # 1. Process output directories -----------------------------------------------------------------
-# The test directories
-files <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = "test")
-
-test_out <- file.path(OUT_DIR, 'test')
-dir.create(test_out)
-
-system2('cp', args = c(paste0(files, collapse = ' '), test_out))
-
-
-
 # Start with experiment 1! And move them into the output direcotry 
-files <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = "exp-1")
-
+files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'exp-1')
 exp1_out <- file.path(OUT_DIR, 'exp-1')
 dir.create(exp1_out)
 
-system2('cp', args = c(paste0(files, collapse = ' '), exp1_out))
+# Start with experiment 1! And move them into the output direcotry 
+files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'test')
+test_out <- file.path(OUT_DIR, 'test')
+dir.create(test_out)
+system2('cp', args = c(paste0(files, collapse = ' '), test_out))
 
+# Proccess the disturbance files
+files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'disturbence-treatments')
+dist_out <- file.path(OUT_DIR, 'disturbence-treatments')
+dir.create(dist_out)
+system2('cp', args = c(paste0(files, collapse = ' '), dist_out))
 
 # 2. Move to local machine ----------------------------------------------------------------------
 message('move the ED outputs with scp')

### 2021-01-14 C.analysis/functions.R

@@ -222,6 +222,39 @@ process_ABG <- function(ed_object, scn_name){
 }
 
 
+process_GPP <- function(ed_object, scn_name){
+  
+  # Make sure that the object being read is ED output. 
+  req_names <- c("basename", "df_scalar", "df_cohort", "df_soil", "df_pft", "outdir")
+  missing_names <- req_names %in% names(ed_object)
+  assert_that(all(missing_names), msg = 'Not an ED object')
+  
+  GPP <- as.data.table(ed_object$df_cohort[[1]])[  ,list(datetime, value = MMEAN_GPP_CO, PFT, DBH, NPLANT)]
+  
+  # Format the data table
+  GPP$year  <- year(GPP$datetime)
+  GPP$month <- month(GPP$datetime)
+  GPP$variable <- "MMEAN_GPP_CO"
+  GPP <- add_pft_names(GPP)
+  GPP <- add_var_info(GPP)
+  GPP$scn <- scn_name 
+  
+  # Convert from a per plant to per area value. 
+  GPP$value <- GPP$NPLANT * GPP$value
+  GPP$unit  <- gsub(pattern = '/pl', replacement = '/m2', GPP$unit)
+  
+  # Change from a per year to a per month flux 
+  GPP$value <- GPP$value/12 
+  GPP$unit  <- gsub(pattern = '/yr', replacement = '', GPP$unit)
+  GPP$description <- gsub(pattern = 'mean - ', replacement = '', GPP$description)
+  
+  return(GPP)
+  
+}
+
+
+
+
 # From that the list from by scn to by variable
 #   Args 
 #     name: the name of the data frame to sort by 
@@ -231,7 +264,8 @@ ed_format_list <- function(name, to_merge){
   
   assert_that(is.list(to_merge))
   assert_that(length(to_merge) >= 1, msg = 'more input files needed to merge')
-  assert_that(name %in% names(to_merge[[1]]))
+  check_names <- sapply(to_merge, function(d){assert_that(name %in% names(d))}) 
+  assert_that(all(check_names))
   
   # Create an empty data table and select name of the list to save 
   out <- data.table()
@@ -242,4 +276,5 @@ ed_format_list <- function(name, to_merge){
   return(out)
   
 }
+
 

### 2021-01-14 C.analysis/plan.R

@@ -6,9 +6,35 @@ source(here::here('C.analysis', 'functions.R'))
 
 ED_OUTPUT_DIR <- here::here('ED-outputs')
 
+# The helper functions that chain together the 
+prep_data <- function(path, scn){
+  
+  assertthat::assert_that(file.exists(path))
+  object <- readRDS(path)
+  
+  data <- list(NPP = process_NPP(object, scn), 
+               LAI = process_LAI(object, scn), 
+               NEP = process_NEP(object, scn), 
+               ABG = process_ABG(object, scn), 
+               GPP = process_GPP(object, scn))
+  
+  return(data)
+}
+format_list <- function(l){
+  
+  assertthat::assert_that(is.list(l))
+  list('NPP' = ed_format_list(name = 'NPP', to_merge = l),
+       'LAI' = ed_format_list(name = 'LAI', to_merge = l), 
+       'NEP' = ed_format_list(name = 'NEP', to_merge = l), 
+       'ABG' = ed_format_list(name = 'ABG', to_merge = l), 
+       'GPP' = ed_format_list(name = 'GPP', to_merge = l))
+}
+
+
+
 # 1. Test Results --------------------------------------
 results_test <- drake_plan(
-
+  
   # Baseline
   test_baseline_ed_object = target(readRDS(file.path(ED_OUTPUT_DIR, "test", "baseline.rds"))), 
   test_baseline_data = target({
@@ -34,56 +60,16 @@ results_test <- drake_plan(
 # 2. Exp 1 ------------------------------------------------------------------------------
 results_exp1 <- drake_plan(
   
-  # Baseline 
-  harvest_0_object =  target(readRDS(file.path(ED_OUTPUT_DIR, "test", "baseline.rds"))), 
-  harvest_0_data = target({
-    scn_name <- 'harvest_0'
-    data <- list(NPP = process_NPP(harvest_0_object, scn_name), 
-                 LAI = process_LAI(harvest_0_object, scn_name), 
-                 NEP = process_NEP(harvest_0_object, scn_name), 
-                 ABG = process_ABG(harvest_0_object, scn_name))
-    data
-  }), 
+  DIR = file.path(ED_OUTPUT_DIR, 'exp-1'), 
   
-  # 45 % FoRTE Treatment 
-  harvest_45_object = target(readRDS(file.path(ED_OUTPUT_DIR, "exp-1", "harvest_45.rds"))), 
-  harvest_45_data = target({
-    scn_name <- 'harvest_45'
[…]

### 2021-01-14 C.analysis/markdowns/Experiment_1.Rmd

@@
     toc: true
-    toc_float: true
+    toc_float:
+      toc_collapsed: true
+    toc_depth: 3
+    theme: lumen
 ---
@@
 
-In experiment 1 we apply the different FoRTE girdling treatments to the ED forest stand.
-This stand starts from near bare ground starting in 1900.
-After a series of tests, we had to exclude pines from the ED configuration.
-
-See [GitHub issue](https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77#issue-687269729) for some more details.
-
-Observations
-
-* NPP winter pattern is interesting
-* By 2030 NPP hasn't recovered 
-* There is a small pattern in the Rh
-
-Questions
-
-* Based on discussion with Lisa do we want to take a look at aboveground bark biomass? Would that be something easy to compare with Max's dendrology band measurements? 
-* Would it make sense to report values as difference harvest 45% - harvest 0%? Or the difference from a refernce value for a single year? 
-* Should we use ED output to calculate resillence? 
-* Should we run ED with climatological inputs?
+In experiment 1 we apply the different FoRTE girdling treatments to the ED forest stand starting from a near bare ground run starting in 1900.
 
@@
 # The defaults for the chunks
-knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 3)
+knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12)
 
@@
 library(purrr)
+library(fortedata)
 
 BASE_DIR <- here::here()
-THEME    <- theme_bw(base_size = 15)
+THEME    <- theme_bw(base_size = 30)
+
+ALPHA <- 0.05
+BCK_GRND1 <- '#dcffcc'
+BCK_GRND2 <- '#9fdfcd'
+BCK_GRND3 <- '#baabda'
+DISTURBANCE_COLORS <- c("0" = "#000000", "45" = "#009E73", "65" = "#0072B2", "85" = "#D55E00")
+
+MONTHLY_BACKGROUND <- function(plot){
+  plot + 
+    geom_rect(aes(xmin=ymd('1900-01-01'), xmax = ymd('1979-12-01'),
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
+    geom_rect(aes(xmin=ymd('1980-01-01'), xmax = ymd('2015-12-01'),
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
+    geom_rect(aes(xmin=ymd('2016-01-01'), xmax = ymd('2029-01-01'),
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
+    labs(caption = 'colored background: looped over 1979 - 1978, obs. 1979 - 2016, avg 2008 - 2018') 
+}
+ANNUAL_BACKGROUND <- function(plot){
+  plot + 
+    geom_rect(aes(xmin=1900, xmax = 1979,
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
+    geom_rect(aes(xmin=1979, xmax = 2015,
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
+    geom_rect(aes(xmin=2015, xmax = 2029,
+                  ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
+    labs(caption = 'colored background: looped over 1979 - 1978, obs. 1979 - 2016, avg 2008 - 2018')
+}
+ADD_BACKGROUND <- function(plot){
+
+  if(any(c('date', 'datetime') %in% names(plot$data))){
+    p <- MONTHLY_BACKGROUND(plot)
+  } else if ('year' %in% names(plot$data)){
+    p <- ANNUAL_BACKGROUND(plot)
+  }
+  return(p)
+}
+
+ADD_DISTURBANCE_LINE_MONTH  <- geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2)
+ADD_DISTURBANCE_LINE_ANNUAL <- geom_vline(xintercept = 2019, color = 'grey', size = 2)
+ADD_COLORS <- scale_color_manual(values = DISTURBANCE_COLORS) 
 ```
@@
 ```{r}
-# Note that this is the object created with the drake frame work, there should be some better way to trigger this. 
+# Note that this is the object created with the drake frame work, 
+# there should be some better way to trigger this.
 exp_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', "exp-1.rds"))
-exp_object <- map(exp_object, function(x){ x[year < 2030, ] }) # Note the year 2030 is not a complete run 
+exp_object <- map(exp_object, function(x){ x[year < 2030, ] })
+
+# Update the data for with the disturbance severity
+add_disturbance_severity <- function(list){
+  y <- map(list, function(x){
+    x[data.table(scn = c("harvest 0", "harvest 45", "harvest 65", "harvest 85"), 
+                 `disturbance severity` = c("0", "45", "65", "85")), on = 'scn'] 
+  })
+  return(y)
+}
+
+exp_object <- add_disturbance_severity(exp_object)
 ```
@@
 
+# Meteorological Inputs
+
+> 
+* dlwrf — downward long wave radiation (W m-2)
+* nbdsf — near infrared beam downward solar radiation (W m-2)
+* nddsf — near IR diffuse downward solar radiation (W m-2)
+* vbdsf — visible beam downward solar radiation (W m-2)
+* vddsf — visible diffuse downward solar radiation (W m-2)
+* prate — precipitation rate (kgH2O m-2 s-1)
+* pres — atmospheric pressure (Pa)
+* hgt — geopotential height (m)
+* ugrd — zonal wind (m s-1)
+* vgrd — meridional wind (m s-1)
+* sh — specific humidity (kgH2O kgAir-1)
+* tmp — air temperature (K)
+
+```{r, fig.width = 12, fig.height = 6}
+met_data <- as.data.table(read.csv(file.path(BASE_DIR, 'A.inputs', 'monthly_mean_met.csv')))
+met_data$date <- ymd(paste(met_data$year, met_data$month, '01', sep = '/'))
+
+long_met_data <- melt(met_data, measure.vars = c("nbdsf", "nddsf", "vbdsf", "vddsf",
+                                                 "prate", "dlwrf", "pres", "hgt", "ugrd", 
+                                                 "vgrd", "sh", "tmp"),
+                      variable.name = 'variable', value.name = 'value')
+long_met_data <- as.data.table(long_met_data[, c("date", "variable", "value", "year")])
+
+var_info <- data.table(variable = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
+                                    'vddsf', 'prate', 'pres', 'hgt', 
+                                    'ugrd', 'vgrd', 'sh', 'tmp'), 
+                       description = c('downward long wave radiation', 
+                                       'near infrared beam downward solar radiation', 
+                                       'near IR diffuse downward solar radiation', 
+                                       'visible beam downward solar radiation',
+                                       'visible diffuse downward solar radiation', 
+                                       'precipitation rate', 
+                                       'atmospheric pressure', 
+                                       'geopotential height', 
+                                       'zonal wind',
+                                       'meridional wind', 
+                                       'specific humidity',
+                                       'air temperature'),  
+                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
+                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
+                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K'))
+long_met_data <- long_met_data[var_info, on = 'variable']
+long_met_data <- long_met_data[variable != 'hgt', ] # exclude the geo spatial height on cause it doesn't 
+```
+
+```{r}
+ADD_BACKGROUND(plot = ggplot(data = long_met_data)) +
+  geom_line(aes(date, value)) + 
+  facet_wrap("variable", scales = 'free') + 
+  labs(title = 'Monthly Time Series of ED Met Data Inputs', y = NULL) + 
+  THEME
+```
+
+Because of the variability in the seasonal cycle is rather large and hides longer trem trends. Let's take a look at the annual min, max and mean.
+
+```{r}
+annual_data <- long_met_data[ , list('value' = mean(value), 'min' = min(value), 'max' = max(value)),
+                              by = list(year, variable)]
+
+ADD_BACKGROUND(plot = ggplot(data = annual_data)) +
+  geom_ribbon(aes(year, ymin = min, ymax = max), fill = 'grey') + 
+  geom_line(aes(year, min), color = 'grey') + 
+  geom_line(aes(year, max), color = 'grey') + 
+  geom_line(aes(year, value)) +
+  facet_wrap('variable', scales = 'free') + 
+  labs(y = NULL) 
+```
+
+Zooming in what does the actual observational data look like?
+
+```{r}
+ggplot(data = annual_data[year %in% 1979:2008, ]) +
+  geom_ribbon(aes(year, ymin = min, ymax = max), fill = 'grey') + 
+  geom_line(aes(year, min), color = 'grey') + 
+  geom_line(aes(year, max), color = 'grey') + 
+  geom_line(aes(year, value)) +
+  facet_wrap('variable', scales = 'free') + 
+  THEME + 
+  labs(y = NULL)
+```
+
 # Baseline Scenario
 
-```{r, fig.width = 3, fig.height = 3}
-ggplot() + 
-  geom_point(data = exp_object$NPP[scn == 'harvest_0', ], aes(year, value)) +
+```{r, fig.width = 12, fig.height = 6}
+# Process the data to plot, the annual values. 
+exp_object$NPP[scn == 'harvest 0', ] %>%  
+    .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
+    .[ , list(year, variable, unit, scn, value)] %>% 
+  unique() -> 
+  NPP_to_plot
+
+exp_object$LAI[scn == 'harvest 0', ] %>%
+  .[  , value := sum(value), by = c('datetime', 'variable', 'unit', 'scn')] %>% 
+  .[ , list(datetime, variable, unit, scn, value)] %>%  
+  unique() -> 
+  LAI_to_plot
+
+exp_object$NEP[scn == 'harvest 0', ] %>%  
+  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
+  .[ , list(year, variable, unit, scn, value)] %>%  
+  unique() -> 
+  NEP_to_plot
+
+exp_object$ABG[scn == 'harvest 0', ] %>%  
+  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
+  .[ , list(year, variable, unit, scn, value)] %>%  
+  unique() -> 
+  ABG_to_plot
+
+exp_object$GPP[scn == "harvest 0", ] %>%  
+  .[  , value := sum(value), by = c('year', 'variable', 'unit', 'scn')] %>% 
+  .[ , list(year, variable, unit, scn, value)] %>%  
+  unique() -> 
+  GPP_to_plot
+
+ADD_BACKGROUND(plot = ggplot(data = NPP_to_plot)) + 
+  geom_line(aes(year, value), size = 1) +
   THEME +
@@
        x = 'Year', 
-       y = unique(exp_object$NPP[scn == 'harvest_0', ][['unit']])) -> 
+       y = unique(exp_object$NPP[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
   NPP_plot
 
-ggplot() + 
-  geom_point(data = exp_object$LAI[scn == 'harvest_0', ], aes(datetime, value)) + 
+ADD_BACKGROUND(plot = ggplot(data = LAI_to_plot)) + 
+  geom_point(aes(datetime, value)) + 
   THEME + 
@@
        x = 'Year', 
-       y = unique(exp_object$LAI[scn == 'harvest_0', ][['unit']])) -> 
+       y = unique(exp_object$LAI[scn == 'harvest 0', ][['unit']]), caption = NULL, size = 0.01) -> 
   LAI_plot
 
-ggplot() + 
-  geom_point(data = exp_object$NEP[scn == 'harvest_0', ], aes(year, value)) +
+ADD_BACKGROUND(ggplot(NEP_to_plot)) + 
+  geom_line(aes(year, value), size = 1) +
   THEME + 
@@
        x = 'Year', 
-       y = unique(exp_object$NEP[scn == 'harvest_0', ][['unit']])) -> 
+       y = unique(exp_object$NEP[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
   NEP_plot
 
-ggplot() + 
-  geom_point(data = exp_object$ABG[scn == 'harvest_0', ], aes(year, value)) + 
+ADD_BACKGROUND(ggplot(ABG_to_plot)) + 
+  geom_line(aes(year, value), size = 1) + 
   THEME + 
@@
        x = 'Year', 
-       y = unique(exp_object$ABG[scn == 'harvest_0', ][['unit']])) -> 
+       y = unique(exp_object$ABG[scn == 'harvest 0', ][['unit']]), caption = NULL) -> 
   ABG_plot
 
-plot_grid(ABG_plot, NPP_plot, NEP_plot, LAI_plot)
+ADD_BACKGROUND(ggplot(GPP_to_plot)) + 
+  geom_point(aes(year, value), size = 1) + 
+  THEME + 
+  labs(title = 'GPP', 
+       x = 'Year', 
+       y = 'kg/m2/yr') -> 
+  GPP_plot
 
+plot_grid(ABG_plot, NPP_plot, NEP_plot, LAI_plot, GPP_plot)
 ```
 
-The base line scenario looks good to me it is intresting just how much smoother the output looks when we swtich to the climatological record instead of weather, do we want to only use the climatological values.
+One potential idea would be to use an ensemble of climate data to create a an envelope of runs.
 
@@
 
-So the differences in total above ground biomass between the different sencaiors is faily small and is really only visible when we zoom in on the disturbance and recovery.
-What is intresting is that the as we would expect the no harvest scenario has a larger biomass until September and then the AGB starts to even out, is pretty similar during winter but then with the next growing season the patches with the more extreme treatments have a larger biomass.
-
-```{r}
+```{r, fig.width = 12, fig.height = 6}
 exp_object$ABG$datetime <- date(exp_object$ABG$datetime)
-data <- exp_object$ABG[  , value := sum(value), by = c("datetime", "scn", "unit", "year")]
-
-data %>%  
-  ggplot() + 
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_point(aes(datetime, value, color = scn, group = interaction(scn, pft_name))) + 
+data <- exp_object$ABG[  , value := sum(value), 
+                         by = c("datetime", "scn", "unit", "year", "disturbance severity")] 
+ggplot(data = data) + 
+  ADD_DISTURBANCE_LINE_MONTH + 
+  geom_line(aes(datetime, value, color = `disturbance severity`, 
+                group = interaction(scn, pft_name)),
+            size = 1.5) + 
   THEME + 
-  labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
-```
-
-```{r}
-data[year %in% 2018:2020 & scn != 'harvest_0', ] %>% 
-  .[datetime >= date('2019-05-01')] %>%  
-  ggplot() + 
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_line(aes(datetime, value, color = scn, group = interaction(scn, pft_name)), size = 1) + 
-  THEME + 
+  ADD_COLORS + 
   labs(x = 'Year', y = unique(exp_object$ABG$unit), title = 'ABG') 
@@
 
-```{r}
-exp_object$NPP %>% 
-  .[year %in% 2018:20205, ] %>% 
-  ggplot() + 
-  geom_vline(aes(xintercept = 2019), color ='black') + 
-  geom_line(aes(year, value, color = scn, group = interaction(scn, pft_name))) + 
+```{r,  fig.width = 12, fig.height = 6}
+ggplot(exp_object$NPP[year %in% 2000:2030, ]) +
+  ADD_DISTURBANCE_LINE_ANNUAL +
+  geom_line(aes(year, value, color = `disturbance severity`, 
+                group = interaction(scn, pft_name)), size = 1.5) + 
   THEME + 
-  labs(x = 'Year', y = unique(exp_object$NPP$unit), title = 'Annual NPP')
+  coord_cartesian(xlim = c(2000, 2025)) + 
+  labs(x = 'Year', y = unique(exp_object$NPP$unit), title = 'Annual NPP') + 
+  ADD_COLORS
 ```
@@
 ```{r}
-h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
-h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
-h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
-h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))
+h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_0_above.rds'))
+h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45_above.rds'))
+h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65_above.rds'))
+h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85_above.rds'))
 
@@
 
-h0$scn  <- 'harvest_0'
-h45$scn <- 'harvest_45'
-h65$scn <- 'harvest_65'
-h85$scn <- 'harvest_85'
+h0$severity  <- '0'
+h45$severity <- '45'
+h65$severity <- '65'
+h85$severity <- '85'
 
@@
 NPP$unit  <- unique(exp_object$NPP$unit)
-NPP <- NPP[, c("datetime", "PFT", "scn", "value", 'unit')]
+NPP <- NPP[, c("datetime", "PFT", "severity", "value", 'unit')]
 NPP$year <- lubridate::year(NPP$datetime)
 NPP$datetime <- date(NPP$datetime)
-NPP_monthly <- NPP 
+monthly_NPP <- NPP
 ```
@@
 ```{r}
-NPP_monthly <- NPP_monthly[, value := sum(value), by = c("datetime", "scn", "unit", "year")]
+NPP_monthly <- monthly_NPP[, value := sum(value), by = c("datetime", "severity", "unit", "year")]
+NPP_monthly$value <- NPP_monthly$value / 12
 
-NPP_monthly %>% 
-  ggplot(aes(datetime, value, color = scn)) + 
-  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_line() + 
+ggplot(NPP_monthly) + 
+  ADD_DISTURBANCE_LINE_MONTH + 
+  geom_line(aes(datetime, value, color = severity)) + 
   THEME + 
-  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
+  ADD_COLORS +
+  labs(title = 'NPP', y = 'kgC/m2/month', x = 'Time') 
 ```
@@
 NPP_monthly %>% 
-  .[year %in% 2018:2030] %>%  
-  ggplot(aes(datetime, value, color = scn)) + 
-  # Note because of the fact that the results are recorded monthly the disturbance date is May 5th.
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_line() + 
+  .[year %in% 2018:2022] %>%  
+  ggplot(aes(datetime, value, color = severity)) + 
+  ADD_DISTURBANCE_LINE_MONTH +
+  ADD_COLORS +
+  geom_line(size = 1.5) + 
   THEME + 
-  labs(title = 'Monthly NPP', y = unique(to_plot_NPP$unit), x = 'Time') 
+  labs(title = 'Monthly NPP', y = 'kgC/m2/month', x = 'Time') 
 ```
-
-Would it make sense to look at the differrence between the baseline and treatment scenarios?
-
-Observations
-
-* Even after 12 years the stand has not recovered in term of NPP values 
-* Also check out the difference in the winter NPP! I wasn't expecting that
 
@@
 exp_object$NEP %>% 
-  dplyr::filter(year >= 2017 & year <= 2021) %>% 
+  dplyr::filter(year >= 2000 & year <= 2021) %>% 
   ggplot() + 
-  geom_vline(xintercept = 2019, color = 'grey', size = 2) + 
-  geom_line(aes(year, value, color = scn)) + 
+  geom_vline(xintercept = 2019, color = 'grey', size = 1.5) + 
+  geom_line(aes(year, value, color = `disturbance severity`), size = 1.5) + 
   THEME + 
+  ADD_COLORS +
   labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Annual NEP')
@@
 ```{r}
-h0_object  <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'test', 'baseline.rds'))
-h45_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_45.rds'))
-h65_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_65.rds'))
-h85_object <- readRDS(file.path(BASE_DIR, 'ED-outputs', 'exp-1', 'harvest_85.rds'))
-
 h0 <- as.data.table(h0_object$df_scalar[[1]])[ ,list(datetime, MMEAN_NEP_PY)]
@@
 NEP$unit  <- unique(exp_object$NEP$unit)
-NEP <- NEP[, c("datetime", "scn", "value" = "MMEAN_NEP_PY")]
-names(NEP) <- c("datetime", "scn", "value")
+NEP <- NEP[, c("datetime", "severity", "value" = "MMEAN_NEP_PY")]
+names(NEP) <- c("datetime", "severity", "value")
 NEP$year <- lubridate::year(NEP$datetime)
@@
   dplyr::filter(year >= 2019 & year <= 2021) %>% 
+  dplyr::mutate(value = value / 12, 
+                units = 'kg/m2/month') %>% 
   ggplot() + 
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_line(aes(datetime, value, color = scn)) + 
+  ADD_DISTURBANCE_LINE_MONTH + 
+  geom_line(aes(datetime, value, color = severity), size = 1) + 
   THEME + 
-  labs(x = 'Year', y = unique(exp_object$NEP$unit), title = 'Monthly NEP')
+  ADD_COLORS +
+  labs(x = 'Year', y ='kg/m2/month', title = 'Monthly NEP')
 ```
@@
 exp_object$LAI %>% 
-  dplyr::filter(year >= 2019 & year <= 2021) %>% 
+  dplyr::filter(year >= 2018 & year <= 2025) %>% 
   ggplot() + 
-  geom_vline(xintercept = date('2019-05-01'), color = 'grey', size = 2) + 
-  geom_line(aes(datetime, value, color = scn)) + 
+  ADD_DISTURBANCE_LINE_MONTH + 
+  geom_line(aes(datetime, value, color = `disturbance severity`)) + 
+  ADD_COLORS +
   THEME + 
@@
 RH %>% 
-ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
+ggplot(aes(datetime, MMEAN_RH_PY, color = severity)) + 
   geom_line() + 
@@
        y = units,
-       x = 'Date Time')
+       x = 'Date Time') + 
+  ADD_COLORS
 ```
@@
 ```{r}
-RH[ , year := year(datetime)] %>% 
-  .[year %in% 2018:2020, ] %>% 
-  ggplot(aes(datetime, MMEAN_RH_PY, color = scn)) + 
-  geom_line() + 
+RH[datetime >= ymd('2018-06-01') & datetime <= ymd('2021-01-01'), ] %>% 
+ggplot(aes(datetime, MMEAN_RH_PY, color = severity)) + 
+  geom_line(size = 1.5) + 
   THEME + 
@@
        y = units,
-       x = 'Date Time')
+       x = 'Date Time') + 
+  ADD_COLORS
 ```
 
-## Mortality
-
-Honestly, I don't really know what the morality output is so might not be easy to interpret but it looks like by 2020 the morality related to the disturbance events has ended.
+## GPP
 
 ```{r}
-h0 <- as.data.table(h0_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
-h45 <- as.data.table(h45_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
-h65 <- as.data.table(h65_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
-h85 <- as.data.table(h85_object$df_cohort[[1]])[  ,list(datetime, mmean_mort_rate_co_carbon)]
+GPP <- exp_object$GPP
+GPP$severity <- GPP$`disturbance severity`
 
-h0$scn  <- 'harvest_0'
-h45$scn <- 'harvest_45'
-h65$scn <- 'harvest_65'
-h85$scn <- 'harvest_85'
+GPP %>% 
+    .[  , value := mean(value), by = c('year', 'variable', 'unit', 'scn', 
+                                      'PFT', 'description', 'severity', 
+                                      'pft_name')] %>% 
+    .[ , list(year, variable, scn, value, unit, pft_name,
+              PFT, description, severity)] -> 
+  GPP_to_plot
 
-mortality    <- rbind(h0, h45, h65, h85)
-units <- '1/yr'
+GPP_to_plot %>%    
+  ggplot(aes(year, value, color = severity)) + 
+  geom_point(size = 3) + 
+  ADD_COLORS + 
+  THEME + 
+  labs(title = 'Annual GPP per Treatment', 
+       y = 'kgC/m2/yr')
+```
 
-mortality[ ,year := year(datetime)] %>% 
-  .[year %in% 2018:2021, ] %>% 
-ggplot(aes(datetime, mmean_mort_rate_co_carbon, color = scn)) + 
-  geom_line() + 
+```{r}
+GPP <- exp_object$GPP
+GPP$severity <- GPP[['disturbance severity']]
+GPP$datetime <- date(GPP$datetime)
+GPP$PFT <- as.character(GPP$PFT)
+
+# Add the cohort labels to the data
+split(GPP, interaction(GPP$datetime, GPP$severity, GPP$PFT)) %>% 
+  lapply(function(x){
+
+    x$cohort <- LETTERS[1:nrow(x)]
+    return(x)
+
+  }) %>%  
+  rlist::list.rbind() -> 
+  GPP_cohort
+
+to_plot <- GPP_cohort[year %in% 2018:2020, ] 
+to_plot$value <- to_plot$value / 12
+to_plot %>%  
+  ggplot(aes(datetime, value, group = interaction(cohort, severity), color = severity)) + 
+  geom_line(size = 2) + 
+  facet_grid(. ~ pft_name) + 
   THEME + 
-  labs(title = 'Monthly Mean Mortlaity Rate per Cohort', 
-       y = units,
-       x = 'Date Time')
+  ADD_COLORS + 
+  labs(title = 'Cohort GPP', 
+       y = 'kgC/m2/month', 
+       x = 'Date')
+```
+
+## Carbon Cycling Resistance
+
+Carbon Cycling Resistance
+
+$$r = \ln \left( {\frac{{F_{{\text{dist}}} }}{{F_{{\text{cont}}} }}} \right)$$ Equation 1 from Gough 2020.
+
+```{r}
+
+exp_object$NPP[year %in% 2018:2030, ] %>%  
+  .[, list(pft_name, year, variable, unit, value, severity = `disturbance severity`)] %>% 
+  # Make the table wide by disturbance severity treatment. 
+  dcast(data = ., pft_name + year + variable + unit ~ severity) %>%
+  # Now gather all of the distubrance treatments into a single column so that the 
+  # control values are recorded in their own column. 
+  melt(data = ., id.vars = c("pft_name", "year", "variable", "unit", "0"), 
+       measure.vars = c('45', '65', '85'), variable.name = 'severity', 
+       value.name = 'treatment_value') %>%  
+  # Rename the columns 
+  .[ , list(pft_name, year, variable, unit, control = `0`, severity, dist = treatment_value)] %>% 
+  # Now calculate r 
+  .[ , r := log(dist/control)] -> 
+  NPP_r
+
+NPP_r %>%   
+ggplot(aes(year, r, color = severity)) + 
+  geom_line(size = 1.5) + 
+  ADD_COLORS + 
+  THEME + 
+  labs(title = 'NPP Resistance Values', 
+       y = expression('r'[npp]), 
+       x = 'Year')
 ```

### 2021-03-03 A.inputs/1.Exp-1_met.R
Version control the met ensemble

@@ -1,51 +1,0 @@
-## FoRTE ED Disturbence: Experiment  1 
-## Bare-ground, 100 year run to mature stand, 0-45-65-85 harvest of C storage pool, 
-## run for 10 more years, using climatology (average of most recent decade), ED default parameters
-## monthly NPP, NEE, LAI, biomass by cohort later we'll want to run multiple times to average out randomness
-## See https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77
-
-# 0. Set Up  ------------------------------------------------------------------------------------
-# Define directories. 
-BASE_DIR  <- here::here()
-WRITE_TO  <- file.path(BASE_DIR, 'exp-1_met2'); dir.create(WRITE_TO, showWarnings = FALSE, recursive = TRUE) # This is the directory to write where the different cases should be set up. 
-EVENT_DIR <- file.path(BASE_DIR, 'A.inputs', 'events')
-
-# Load the required pacakges and functions. 
-to_source <- file.path(BASE_DIR, "B.ed", "0A.dependencies_functions.R")
-stopifnot(file.exists(to_source))
-source(to_source)
-
-
-# 1. Define ED Runs ------------------------------------------------------------------------------------
-# Use the case data frame to set up the different runs with the 0-45-65-85 harvest of the C storage pool. 
-
-# Start by defining the ED run information. 
-IYEARA <- 1900
-IYEARZ <- 2034
-
-# Find the event files
-event_file <- list.files(EVENT_DIR, pattern = '1day_above.xml', full.names = TRUE)
-cases      <- gsub(pattern = '.xml', replacement = '', x = basename(event_file))
-assertthat::assert_that(all(file.exists(event_file)))
-# Important! When reading in the event_file paths they must be written out as a string, 
-# otherwise ED will fail but will throw an uninformative error message realted to 
-# missing a different ED input. 
-event_file <- paste0("'", event_file, "'") 
-
-# TODO I had to manuual change the ED met dirver pointer, this shoudl be something that can be changed with a point in a direction! 
-# Save all of the ED information as a single data frame. 
-case <- data.frame(casename = cases, IYEARA = IYEARA, IYEARZ = IYEARZ,
-                   EVENT_FILE = event_file, 
-                   ED_MET_DRIVER_DB = file.path(INPUT_DIR, "NARR-ED2_met3", "ED_MET_DRIVER_HEADER"))
-
-
-
[…]

### 2021-03-03 A.inputs/1.generate_met_ensemble.R

@@ -19,14 +19,15 @@ library(hdf5r)
 # These files have been committed to this repo for purposes of reproducibility. 
 # TODO is this legal or no? 
 BASE_DIR  <- here::here()
-PECAN_DIR <- '/people/dorh012/forte-workflow/pecan'
-assertthat::assert_that(dir.exists(PECAN_DIR))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'read_ed2in.R'))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'write_ed2in.R'))
+#PECAN_DIR <- '/people/dorh012/forte-workflow/pecan'
+#assertthat::assert_that(dir.exists(PECAN_DIR))
+#source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'read_ed2in.R'))
+#source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'write_ed2in.R'))
 
 # Define the directories
 BASE_DIR <- here::here()
-MET_DIR <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
+MET_DIR  <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
+WRITE_TO <- here::here("ED-outputs", "met_results")
 
 # Make sure that the met data dir exists. 
 assertthat::assert_that(dir.exists(MET_DIR))
@@ -55,13 +56,36 @@ prep_dir <- function(out_dir){
 
 # TODO there should probably be some way to itterate over this 
 # -----------------------------------------------------------------------------------------------
-met2_dir <- here::here(BASE_DIR, 'A.inputs', 'NARR-ED2-met2')
-dir.create(met2_dir, showWarnings = FALSE)
 
-# Prepare the dir and then generate the future met data as some average of data over several years.
-prep_dir(met2_dir)
-generate_future_met(out_dir = met2_dir, use_yrs = 2000:2005, startYr = 2000, endYr = 2035)
+complete_years <- 1980:2008 
 
+start <- seq(from = min(complete_years), by = 5,length.out = 8)
+end   <- start + 4
+
+periods <- mapply(function(x,y){x:y}, x = start, y = end, SIMPLIFY = FALSE)
+names(periods) <- paste0('met', 1:length(periods))
+
+
+for(m in names(periods)){
+  
+  main_dir <- here::here(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
[…]

### 2021-03-03 A.inputs/1A.met_input_processing.R

@@ -1,104 +1,0 @@
-# This script only needs and should only be run once. After the met data is generated 
-# there is no need to re run it. See READ ME before running. 
-#
-# It is fairly hacky but cdo/netcdf/ncdf4 won't can't be used to write the h5 file out. 
-# 
-# 0. Set Up -----------------------------------------------------------------------------------------
-# Load the libraries 
-library(magrittr)
-library(here)
-library(data.table)
-library(ncdf4)
-library(purrr)
-library(lubridate)
-library(assertthat)
-library(hdf5r) # Unfortunately this package cannot be installed on pic the met files must be devvloped on 
-# local machine then transfered to pic.
-
-# There were some issues installing pecan specfic scripts should be sourced from the pecan projects. 
-# See https://github.com/PecanProject/pecan to downlaod peacn 
-# git show-branch 
-# [develop] Merge pull request #2570 from infotroph/book-headers
-# commit b264ce700a1892726a7e4ec21c3187152297a190 
-# brach devlop
-# These files have been committed to this repo for purposes of reproducibility. 
-# TODO is this legal or no? 
-BASE_DIR  <- here::here()
-PECAN_DIR <- '/people/dorh012/forte-workflow/pecan'
-assertthat::assert_that(dir.exists(PECAN_DIR))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'read_ed2in.R'))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'write_ed2in.R'))
-
-# Define the directories
-BASE_DIR <- here::here()
-MET_DIR <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
-METOUT_DIR <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2')
-dir.create(METOUT_DIR, showWarnings = FALSE)
-
-# Make sure that the met data dir exists. 
-assertthat::assert_that(dir.exists(MET_DIR))
-
-source(here::here('A.inputs', 'helper_functions.R'))
-
-# -----------------------------------------------------------------------
-# Start by creating a new met header file 
-new_header <- modify_met_header(to = METOUT_DIR)
-
-# Copy over all of the files exist in the origial copy into the new location 
-files     <- list.files(MET_DIR, '*.h5', full.names = TRUE)
-new_files <- copy_files_func(files = files, new_dir = METOUT_DIR)
-
[…]

### 2021-03-03 A.inputs/helper_functions.R

@@ -44,7 +44,10 @@ modify_met_header <- function(to){
   metdriver_file <- file.path(here::here('A.inputs', 'NARR-ED2-main'), 'ED_MET_DRIVER_HEADER')
   met_file       <- readLines(metdriver_file)
   met_file[[1]]  <- "Modified for the forte-disturbance project"
-  met_file[[3]]  <- paste0(to, '/')
+  
+  bd <- basename(to)
+  pic_dir <- '/people/dorh012/forte-disturbance/A.inputs'
+  met_file[[3]]  <- paste0(pic_dir, '/', bd, '/')
   
   out_file <- file.path(to, 'ED_MET_DRIVER_HEADER')
   writeLines(met_file, out_file)
@@ -209,7 +212,7 @@ generate_future_met <- function(out_dir, use_yrs, startYr, endYr){
   
   # Check the inputs 
   assert_that(dir.exists(out_dir))
-  assert_that(all(is.integer(use_yrs), length(use_yrs) >= 2))  
+  assert_that(all(is.integer(use_yrs)))  
   assert_that(startYr < endYr)
   
   # Figure out which files we do have 
@@ -251,15 +254,17 @@ generate_future_met <- function(out_dir, use_yrs, startYr, endYr){
       # Caculate the daily average value for each variable 
       df <- get_nc_mean(ncs)
       
+      
       # This code was based off of https://github.com/PecanProject/pecan/blob/3b0cb56bba1eb90d1f7bbe3b0da3510f681519ba/models/ed/R/met2model.ED2.R
       # and requires hdf5 dependencies (might not work on pic). 
       temp_out <- file.path(paste0(mon, '.h5')) 
       file.copy(ncs[[1]], temp_out)
+      fill_in_h5(file = temp_out, df = df, n_steps = n_steps)
       
       
       # Now copy over the netcdf files to all of the files.
       new_files <- file.path(out_dir, paste0(2010:endYr, mon, '.h5'))
-      file.copy(temp_out, new_files)
+      file.copy(temp_out, new_files, overwrite = TRUE)
       
       # Remove the temp file
       file.remove(temp_out)
@@ -341,3 +346,98 @@ fill_in_h5 <- function(file, df, n_steps){
   return(file)
   
 }
+
+
+# Extract data from the h5 files and save as csv files 
+# 
+# Args 
+#   DIR: str directory path that contains the different met data 
+#   met: str name of the specific meteorology dir to process 
[…]

### 2021-03-03 A.inputs/1B.met_input_ensemble.R
clean up'

@@ -1,58 +1,0 @@
-# Generate the ensemble of climate runs 
-
-
-# So there really should be a set of unchanged met data, that is dowlaoded 
-# bcause everytime we manipualte it we loose the data in the recent history that was
-# use :( 
-
-# Also because we look the climatological average over this period in time I think we are looking at the 
-# ensemble mean hmmmmm hmmmm would it be as simple as shifting the runs. 
-
-# bleh it is really confusing and i dont wannt tho think about this at all!
-
-
-
-
-# 0. Set Up -----------------------------------------------------------------------------------------
-# Load the libraries 
-library(magrittr)
-library(here)
-library(data.table)
-library(ncdf4)
-library(purrr)
-library(ncdf4)
-library(lubridate)
-
-
-# There were some issues installing pecan specfic scripts should be sourced from the pecan projects. 
-# See https://github.com/PecanProject/pecan to downlaod peacn 
-# git show-branch 
-# [develop] Merge pull request #2570 from infotroph/book-headers
-# commit b264ce700a1892726a7e4ec21c3187152297a190 
-# brach devlop
-# These files have been committed to this repo for purposes of reproducibility. 
-# TODO is this legal or no? 
-BASE_DIR  <- here::here()
-PECAN_DIR <- '/people/dorh012/forte-workflow/pecan'
-assertthat::assert_that(dir.exists(PECAN_DIR))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'read_ed2in.R'))
-source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'write_ed2in.R'))
-
-# Define the directories
-BASE_DIR <- here::here()
-MET_DIR <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2')
-
-# Make sure that the met data dir exists and that there is data for the last met year. 
-last_met_year <- 2030
-assertthat::assert_that(dir.exists(MET_DIR))
-
-# 1. Find Met Inputs -----------------------------------------------------------------------
-h5_file_names <- list.files(MET_DIR, '*.h5')
-assertthat::assert_that(any(grepl(pattern = last_met_year, h5_file_names)))
-
-
-
-
-
-
-

### 2021-03-03 A.inputs/1C.produce_met_csv.R

@@ -1,121 +1,0 @@
-# Process all of the met h5 files and save monthly and annual csv files. 
-# 0. Set Up ----------------------------------------------------------------------------------
-library(lubridate)
-library(magrittr)
-library(data.table)
-library(ncdf4)
-library(assertthat)
-
-WRITE_TO <- '/qfs/people/dorh012/forte-disturbance/ED-outputs/met_results'
-
-# Extract data from the h5 files and save as csv files 
-# 
-# Args 
-#   DIR: str directory path that contains the different met data 
-#   met: str name of the specific meteorology dir to process 
-# Return: str vector of the csv files, located in a temporary directory 
-extract_met_data <- function(DIR, met_dir){
-  
-  dir <- file.path(DIR, paste0('NARR-ED2_', met_dir))
-  
-  header_file <- list.files(dir, 'HEADER', full.names = TRUE)
-  header_contents <- readLines(header_file)
-  vars <- unlist(strsplit(header_contents[[6]], split = ' '))
-  vars <- vars[vars != 'co2']
-  
-  h5_files <- list.files(dir, '.h5', full.names = TRUE)
-  
-  results <- c()
-  temp_dir <- tempdir()
-  
-  for(path in h5_files){
-    
-    nc <- nc_open(path)
-    values <- sapply(X = vars, FUN = ncvar_get, nc = nc, USE.NAMES = TRUE)
-    values <- as.data.frame(values)
-    
-    year_month <- gsub(basename(path), pattern = '.h5', replacement = '')
-    year <- as.integer(substr(year_month, 1, 4))
-    month <- gsub(pattern = year, replacement = '', x = year_month)
-    month_list <- toupper(month.abb)
-    month <-  which(month == month_list)
-    
-    out      <- cbind(year = year, month = month,  t_index =  1:nrow(values), values)
-    out_file <- file.path(temp_dir, paste0(basename(path),'.csv'))
-    write.csv(out, file = out_file, row.names = FALSE)
-    results <- c(results, out_file)
-    
-  }
-  
-  return(results)
-  
-}
-
-
-# Calculate the monthly values from the met data files 
-# Args 
-#   files: str vector path to the files  
-# Return: a data frame of the monthly values
-get_monthly_mean <- function(files){
-  
-  # Check the inputs 
-  assert_that(all(file.exists(files)))
[…]

### 2021-03-03 B.ed/0A.dependencies_functions.R
set up downstream scripts to process the met outputs

@@ -127,7 +127,7 @@ setup_ed_run <- function(case, input_dir = INPUT_DIR, write_to = WRITE_TO){
     
     LU_DATABASE = file.path(INPUT_DIR, "EDI", "ed_inputs", "glu"),
     THSUMS_DATABASE = file.path(INPUT_DIR, "EDI", "ed_inputs/"),
-    ED_MET_DRIVER_DB = file.path(INPUT_DIR, "NARR-ED2", "ED_MET_DRIVER_HEADER"),
+   # ED_MET_DRIVER_DB = file.path(INPUT_DIR, "NARR-ED2", "ED_MET_DRIVER_HEADER"),
     
     # UMBS soil characteristics (from Gough et al. 2010 FEM)
     ISOILFLG = 2,  # A flag that allows the soil characteristics to be defined below. 
@@ -172,14 +172,21 @@ setup_ed_run <- function(case, input_dir = INPUT_DIR, write_to = WRITE_TO){
     from_case_tags <- NULL
   }
   
-  # Combine the tag information, make sure that if there are duplicate entries 
-  # that the values from the case tags is used. 
-  ed2in_tags <- append(ed2in_tags, config_tags)
-  default_to_keep <- base::setdiff(names(ed2in_tags), names(from_case_tags))
-  ed2in_tags <- append(default_to_keep, from_case_tags)
+  # # Combine the tag information, make sure that if there are duplicate entries 
+  # # that the values from the case tags is used. 
+  # ed2in_tags <- append(ed2in_tags, config_tags)
+  # default_to_keep <- base::setdiff(names(ed2in_tags), names(from_case_tags))
+  # ed2in_tags <- append(default_to_keep, from_case_tags)
+  # 
+  # # Modify the ED2IN file with the updated information
+  # ed2in <- modifyList(ed2in_template, ed2in_tags)
+  
+  ed2in_tags <- append(append(ed2in_tags, config_tags), from_case_tags)
   
   # Modify the ED2IN file with the updated information
   ed2in <- modifyList(ed2in_template, ed2in_tags)
+  
+  
   
   # Write the ed2 in file out 
   file <- file.path(OUTOUT, "ED2IN")

### 2021-03-03 B.ed/1.Experiment-1.R

@@ -5,6 +5,8 @@
 ## See https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77
 
 # 0. Set Up  ------------------------------------------------------------------------------------
+library(data.table)
+
 # Define directories. 
 BASE_DIR  <- here::here()
 WRITE_TO  <- file.path(BASE_DIR, 'exp-1'); dir.create(WRITE_TO, showWarnings = FALSE, recursive = TRUE) # This is the directory to write where the different cases should be set up. 
@@ -21,7 +23,7 @@ source(to_source)
 
 # Start by defining the ED run information. 
 IYEARA <- 1900
-IYEARZ <- 2034
+IYEARZ <- 2049
 
 # Find the event files
 event_file <- list.files(EVENT_DIR, pattern = '1day_above.xml', full.names = TRUE)
@@ -36,9 +38,15 @@ event_file <- paste0("'", event_file, "'")
 case <- data.frame(casename = cases, IYEARA = IYEARA, IYEARZ = IYEARZ, EVENT_FILE = event_file)
 
 
-met_headers <- data.table(ED_MET_DRIVER_DB = c(file.path(INPUT_DIR, "NARR-ED2", "ED_MET_DRIVER_HEADER"), 
-                                               file.path(INPUT_DIR, "NARR-ED2", "ED_MET_DRIVER_HEADER_met2")), 
-                          met_name = c('met1', 'met2'))
+met_headers <- data.table(ED_MET_DRIVER_DB = c(file.path(INPUT_DIR, "NARR-ED2_met1", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met2", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met3", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met4", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met5", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met6", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2_met7", "ED_MET_DRIVER_HEADER"), 
[…]

### 2021-03-03 B.ed/2.post_processing.R

@@ -21,15 +21,6 @@ dir.create(exp1_out)
 rds_files <- list.files(file.path(BASE_DIR, 'exp-1'), pattern = '.rds', recursive = TRUE, full.names = TRUE)
 system2('cp', args = c(paste0(rds_files, collapse = ' '), exp1_out))
 
-
-# Process the runs from the disturbance runs, these are all the different disturbance 
-# treatments we applied to ED while we were trying out different configurations. 
-files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'disturbence-treatments')
-dist_out <- file.path(OUT_DIR, 'disturbence-treatments')
-dir.create(dist_out)
-system2('cp', args = c(paste0(files, collapse = ' '), dist_out))
-
-# 2. Move to local machine ----------------------------------------------------------------------
 message('move the ED outputs with scp')
 
 

### 2021-03-03 C.analysis/functions.R
Adding met experiment 1 files to github

@@ -1,280 +1,0 @@
-# Custom functions for the drake workflow that is used to process the 
-# monthly ED output files into data frames that are easy to work with. 
-
-
-# Add the ED pft names of the data table 
-#   Args  
-#     dt: the data table of ED information, it must have PFT column 
-#   Return: a data table with the pft names 
-add_pft_names <- function(dt){
-  
-  # Check the inputs 
-  assert_that(is.data.table(dt))
-  assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
-  
-  # Data table of the plant functional types in ED. 
-  pft_names <- data.table(pft = c(6, 8, 9, 10, 11, 7) , 
-                          pft_name = c('Northern North American pines', 
-                                       'Late-successional North American conifers', 
-                                       'Temperate broadleaf, early successional',  
-                                       'Temperate broadleaf, mid-successional', 
-                                       'Temperate broadleaf, late successional', 
-                                       'Southern North American pines'))
-  
-  
-  # Add the pft names to the pft data table. 
-  named_dt <- na.omit(dt[pft_names, on = c("PFT" = "pft")])
-  return(named_dt)
-  
-}
-
-# KALYN DOING THIS ONE WRONG, IT IS SUPPOSED TO BE THE OTHER WAY 
-# Add the UMBS species and the pft name to a ED data table
-#   Args 
-#     dt: a data table of ED output that contains the PFT column
-# Return: a data table of ED output with the PFT names and UMBS species 
-# UMBS_to_ED <- function(dt){
-#   
-#   # Check the inputs 
-#   assert_that(is.data.table(dt))
-#   assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
-#   
-#   # A Data table that maps the ED PFT to the UMBS species 
-#   mid_success       <- data.table(pft = 10, UMBS = c('ACRU','ACPE',  'FAGR', 'ACSA'), stringsAsFactors = FALSE)
-#   temp_broadl_late  <- data.table(pft = 11, UMBS = c('QURU', 'TCSA'), stringsAsFactors = FALSE)
[…]

### 2021-03-03 C.analysis/exp-1_met.Rmd
Adding met experiment 1 files to github

---
title: 'Exp-1 Met Results'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: true
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_chunk$set(dev = c('pdf', 'png'), 
        fig.align = 'center', fig.height = 5, fig.width = 8.5, 
        pdf.options(encoding = "ISOLatin9.enc"))

library(ggplot2)
library(data.table)
library(fortedata)
library(magrittr)
library(fortedata)
library(ggpmisc)
library(assertthat)
library(GGally)
library(knitr)

# Define the base directory. 
BASE_DIR <- "/Users/dorh012/Documents/2020/FoRTE/forte-disturbance"
FIG_OUT <- file.path(BASE_DIR, "C.analysis", "FIGS_exp-1_met")
dir.create(FIG_OUT, showWarnings = FALSE)

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
ALPHA <- 0.05
BCK_GRND1 <- '#dcffcc'
BCK_GRND2 <- '#9fdfcd'
BCK_GRND3 <- '#baabda'
FORTE_SEVERITY_COLORS <- c("#000000", "#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('0 %', '45 %', '65 %', '85 %')
```

# Objective

Run ED using different climate to create an ensemble of runs

* How does that impact ED results? 
* What impact does it make on dimensions of stability? 
* Is there a relationship between the inputs & metrics? 
* How do we want to calculate recovery? 
  * When it reaches the control? When it reaches a steady state?

# Compare Inputs

```{r}
# Import the meterolical inputs. 
RESULTS_DIR <- file.path(BASE_DIR, "ED-outputs", "met_results")

# Create a data frame of all the meterolgical data used. 
list.files(RESULTS_DIR, pattern = 'yr', full.names = TRUE) %>%  
  lapply(function(x){

    name <- gsub(x = basename(x), pattern = '_yr.csv', replacement = '')
    data <- read.csv(x, stringsAsFactors = FALSE)
    data$scn <- name
    return(data)

  }) %>%  
  dplyr::bind_rows() %>%  
  as.data.table -> 
  data

# A mapping file of variable names, description, and units. 
# This will be used to add more information to the met data data frame. 
var_info <- data.table(variable = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K'), stringsAsFactors=FALSE)

data <- data[var_info, on = 'variable']
data <- na.omit(data)

data$label <- paste0(data$description, ' (', data$units, ')')

```

```{r}
# Define function that generates the standard met input plot 
# Args 
#   data: data table of the met data to plot
#   yr1: the first year of met data to plot, default set to 1950
#   yr2: the final year of met data to plot, default set to 2030 
# Return: a faceted line plot of the met data, ggplot2 object
met_plot <- function(d, yr1 = 1950, yr2 = 2040){

  # Color the background 
  ALPHA <- 0.05
  BCK_GRND1 <- '#dcffcc'
  BCK_GRND2 <- '#9fdfcd'
  BCK_GRND3 <- '#baabda'

  d[year >= yr1 & year <= yr2] %>%  
    ggplot() + 
    geom_rect(aes(xmin=yr1, xmax = 1979,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
    geom_rect(aes(xmin=1979, xmax = 2010,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
    geom_rect(aes(xmin=2010, xmax = yr2,
                  ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
    labs(caption = 'colored background: looped over obs., obs., 5 year obs. mean ') + 
    geom_line(aes(year, value, linetype = scn)) + 
    facet_wrap('label', scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) + 
    THEME + 
    labs(y = NULL, x = 'Year', title = 'Annual Mean Meteorological Inputs') + 
    theme(legend.position = 'none') -> 
    plot

  return(plot)

}

# Split the met data up into two different plots, this will make the plots easier to  look at. 
var_list <- unique(data$variable)
pannels_n1 <- ceiling(length(var_list)/2)
```

Up until 2010 the runs use the same meteorology inputs, starting in 2010, about a decade before the disturbance we start to use different meteorology inputs based on 5 year means sampled from the observational data set.

## Plot Climate Input

```{r}
# Plot the first half of the met input values 
met_1 <- met_plot(data[variable %in% var_list[1:pannels_n1]])
ggsave(met_1, filename = file.path(FIG_OUT, "met_1.png"), height = 5,width = 8.5)
met_1
```

```{r}
# Plot the second half of the met input values 
met_2 <- met_plot(data[variable %in% var_list[pannels_n1+1:length(var_list)]])
ggsave(met_2, filename = file.path(FIG_OUT, "met_2.png"), height = 5,width = 8.5)
met_2
```

The inputs for the two different met records starting in 2010, so that that ED has a steady climate for the ~10 yrs before the disturbance, the met ensemble captures a wide range of different climates.

# ED Results

```{r}
annual_data <- as.data.table(read.csv(file.path(BASE_DIR, "ED-outputs", "results",
                                                "exp-1-yr.csv"), stringsAsFactors = FALSE))
```

## Plot Baseline/Control

```{r}
annual_data$label <- paste0(annual_data$description, " (", annual_data$unit, ")")

annual_data[scn == 'harvest_0_1day_above', ] %>%  
  .[ , severity := '0'] %>% 
  ggplot() + 
  #    geom_rect(aes(xmin=1900, xmax = 1979,
  #                ymin = -Inf, ymax = Inf), fill = BCK_GRND1, alpha = ALPHA) +
  #  geom_rect(aes(xmin=1979, xmax = 2010,
  #                ymin = -Inf, ymax = Inf), fill = BCK_GRND2, alpha = ALPHA) +
  #  geom_rect(aes(xmin=2010, xmax = 2048,
  #                ymin = -Inf, ymax = Inf), fill = BCK_GRND3, alpha = ALPHA) + 
   # geom_vline(xintercept = 2019, color = 'black') + 
  geom_line(aes(year, value, linetype = met)) + 
 facet_wrap('label', scales = 'free',  
            labeller = label_wrap_gen(width = 50, multi_line = TRUE)) + 
  THEME + 
  theme(legend.position = 'none')+
  labs(title = 'Control Runs', 
       y = NULL, 
       x = "Year")  -> 
  control_plot

ggsave(control_plot, filename = file.path(FIG_OUT, "control_plot.png"), height = 5,width = 8.5)

#, 
#       caption = 'vertical line at 2019, when the FoRTE treatment is #applied\ncolored background: looped over obs., obs., 5 year obs. mean ') 
```

Yes changing the meteorology input files does make a difference in the baseline scenarios, as we can tell based on the vertical line by the time the FoRTE treatment is applied the ED forest stands diverged from one another.

# Disturbance

What is the difference between the different treatments and the control (0%) runs?

```{r}
baseline <- annual_data[scn == 'harvest_0_1day_above', ][ , .(year, variable, description, unit, met, baseline = value)]
disturbance <- annual_data[scn != 'harvest_0_1day_above', ]
setnames(x = disturbance, old = "value", new = "disturbance", skip_absent = TRUE)
dt <- baseline[disturbance, on = c( "year", "variable", "description", "unit", "met")]
dt[ , value := disturbance - baseline]
dt[ , severity := gsub(pattern = "harvest_|_1day_above", replacement = "", x = scn)]
```

## Plot Disturbance

```{r}
dt$`Disturbance Severity` <- paste0(dt$severity, ' %')

dt[year >= 2015, ] %>% 
  ggplot(aes(year, value, color = `Disturbance Severity`, linetype = met)) + 
 # geom_vline(xintercept = 2019, color = 'black') + 
  geom_line() + 
   facet_wrap('label', scales = 'free',  
            labeller = label_wrap_gen(width = 30, multi_line = TRUE)) + 
  THEME + 
  labs(y = 'Disturbance - Baseline', 
       x = 'Year') + 
  #, 
  #     caption = 'vertical line 2019 disturbance event')  + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(linetype = FALSE) -> 
  disturbance_plots
ggsave(disturbance_plots, filename = file.path(FIG_OUT, "disturbance_plot.png"), 
       height = 5,width = 8.5)
disturbance_plots
```

The over shoot in the met2 track for the carbon fluxes is really interesting! Do we think that would happen if the met1 run continued longer?

```{r}
# dt$`Disturbance Severity` <- paste0(dt$severity, ' %')
# 
# 
# lapply(unique(dt$label), function(l){
#   
#   dt[year >= 2015, ] %>% 
#   .[label == l, ] %>% 
#    ggplot(aes(year, value, color = `Disturbance Severity`, linetype = met)) + geom_vline(xintercept = 2019, color = 'black') + 
#   geom_line() + 
#    facet_wrap("scn") + 
#   THEME + 
#   theme(  strip.background = element_blank(),
#   strip.text.x = element_blank()) +
#   labs(y = 'Disturbance - Baseline', 
#        x = 'Year') + 
#   #, 
#   #     caption = 'vertical line 2019 disturbance event')  + 
#   scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
#   guides(linetype = FALSE) + 
#   labs(title = l)
#   
# }) -> 
#   plots
# names(plots) <- unique(dt$label)
# 
# 
# out <- plots$`Annual  Net primary productivity (kgC/m2)`
# ggsave(out, filename = file.path(FIG_OUT, "disturbance_NPP.png"), 
#        height = 5,width = 8.5)
# 
# out <- plots$`Annual  Net Ecosystem productivity (kg/m2)`
# ggsave(out, filename = file.path(FIG_OUT, "disturbance_NEP.png"), 
#        height = 5,width = 8.5)
# 
# out <- plots$`Annual  Gross primary productivity (kgC/m2)`
# ggsave(out, filename = file.path(FIG_OUT, "disturbance_GPP.png"), 
#        height = 5,width = 8.5)
# 
# out <- plots$`Annual Mean Above-ground biomass (kgC/m2)`
# ggsave(out, filename = file.path(FIG_OUT, "disturbance_AGB.png"), 
#        height = 5,width = 8.5)

```

# Dimensions of Stability

```{r}
# Find the recovery at specific time places. 
# Args 
#   d: a data frame of the annual ED outputs
#   x: a vectory of the years to calculate the recovery at. 
# Return: a data frame of the recovery values
recovery_at_X <- function(d, x){

# Check inputs 
req_names <- c('scn', 'year', 'variable', 'description', 'unit', 'value', 'met')
assert_that(has_name(d, req_names))

baseline_scn <- "harvest_0_1day_above" 
assert_that(baseline_scn %in% d$scn)

assert_that(all(x %in% d$year))

# Separate the control or baseline scneario values from the output from the other 
# treatment groups. 
baseline_values <- d[scn == baseline_scn][ , c("year", "variable", "value", "met")]
names(baseline_values) <- c("year", "variable", "baseline", "met")
treatment_values <- d[scn != baseline_scn][ , c("scn", "year", "variable", "description", 
                                                "unit", "treatment" = "value", "met", "label")]
names(treatment_values) <- c("scn", "year", "variable", "description", 
                                                "unit", "treatment", "met", "label")
# Combine the control and the treatment values as a wide df, then take the # ln(treatment/control) per year / variable / met realization. 
disturbance_d <- treatment_values[baseline_values, on = c("year", "variable", "met")][year %in% 2019:2050]
disturbance_d[ , value := log(treatment/baseline)]

# Now that the ln(treatment/control) has been calculated we can use these values to calculate
# the various dimensions of ecosystem stability. 
recovery_d <- disturbance_d[year %in% x]
recovery_d <- recovery_d[ , c("scn", "year", "variable", "description", 
                              "unit", "met",  "value")]
names(recovery_d) <- c("scn", "year", "variable", "description", 
                       "unit", "met",  "recovery")
dt <- recovery_d

# Add the disturbance severity column. 
dt$`Disturbance Severity` <- paste0(gsub(x = dt$scn,
                                         pattern = "harvest_|_1day_above", 
                                         replacement = ""), " %")

# Format and apply factor information to the variable name so the dimensions of stability plot 
# step through the carbon cycle in order GPP -> NPP -> NEP. 
dt$variable <- gsub(pattern = "_patch", replacement = "", x = dt$variable)
dt$variable <- factor(x = dt$variable, levels = c("GPP", "NPP", "NEP"), ordered = TRUE)
dt <- na.omit(dt)
return(dt)

}

recovery_at_X_data <- recovery_at_X(annual_data, c(2025, 2030, 2035, 2040))
```

```{r}
# Calculate the various dimensiosn of stability 
# Args
#   d: data table of the annual ED results 
# Return: data table of the dimensions of stability. 
calculate_dimensions_stability <- function(d){

# Check inputs 
req_names <- c('scn', 'year', 'variable', 'description', 'unit', 'value', 'met')
assert_that(has_name(d, req_names))

baseline_scn <- "harvest_0_1day_above" 
assert_that(baseline_scn %in% d$scn)

# Separate the control or baseline scneario values from the output from the other 
# treatment groups. 
baseline_values <- d[scn == baseline_scn][ , c("year", "variable", "value", "met")]
names(baseline_values) <- c("year", "variable", "baseline", "met")
treatment_values <- d[scn != baseline_scn][ , c("scn", "year", "variable", "description", 
                                                "unit", "treatment" = "value", "met", "label")]
names(treatment_values) <- c("scn", "year", "variable", "description", 
                                                "unit", "treatment", "met", "label")
# Combine the control and the treatment values as a wide df, then take the # ln(treatment/control) per year / variable / met realization. 
disturbance_d <- treatment_values[baseline_values, on = c("year", "variable", "met")][year %in% 2019:2050]
disturbance_d[ , value := log(treatment/baseline)]

# Now that the ln(treatment/control) has been calculated we can use these values to calculate
# the various dimensions of ecosystem stability. 
treatment_yr <- 2019 # The year of the girdel

# Resistance: ${r}_{t} = ln(\frac{{F}_{disturbance}}{{F}_{control}})$ at the year year 
# of disturabnce (2019). 
resitance_d <- disturbance_d[year == treatment_yr][ , c("scn", "variable",
                                                        "description", "unit", "met",  "value")]
names(resitance_d) <- c("scn", "variable", "description", "unit", "met",  "resitance")

# Resilience : ${r}_{i} = ln(\frac{{F}_{disturbance}}{{F}_{control}}) = i + {r}_{i} * t$
# Calculate the resilience (the slope of the recovery starting at the treatment_yr)
d_to_use <- disturbance_d[year >= 2019 & year <= 2030] # Do not include data from pre treatment years.

# Get the result from the linear regression per variable / scn / met ensemble member. 
d_list <- split(d_to_use, interaction(d_to_use$variable, d_to_use$scn, d_to_use$met), 
                drop = TRUE) 
fits <- lapply(d_list, function(input){lm(value ~ year, data = input)})

# Get the slope from the fits, this is the resilience value. 
resilience_values <- sapply(fits, function(x){x$coefficients[2]})
resilience_d <- data.table(info = names(resilience_values), resilience = resilience_values)
resilience_d <- resilience_d[, c("variable", "scn", "met", "x") := tstrsplit(info, ".",
                                                                             fixed=TRUE)]
resilience_d <- resilience_d[, c("resilience", "variable", "scn", "met")]

# Now using the fits calculate the temporal stability. 
# ${s}_{t} = \frac{1}{ SE of resid}$
se_resid <- sapply(fits, function(model){sqrt(deviance(model)/df.residual(model))})
temp_stability_d <- data.table(info = names(se_resid), temp_stability = 1/se_resid)
temp_stability_d <- temp_stability_d[, c("variable", "scn", "met") := tstrsplit(info, ".",
                                                                             fixed=TRUE)]
temp_stability_d <- temp_stability_d[, c("temp_stability", "variable", "scn", "met")]

# Finally calculate the recovery, for now we will define recovery at year 2039
# two decades after the year of the girdel. 
recovery_d <- disturbance_d[year == 2030]
recovery_d <- recovery_d[ , c("scn", "variable", "description", "unit", "met",  "value")]
names(recovery_d) <- c("scn", "variable", "description", "unit", "met",  "recovery")
recovery_d$recovery <- recovery_d$recovery

# Combine all dimensions of stability into a single data table. 
resitance_d[resilience_d, on = c("variable", "scn", "met")] %>% 
  .[temp_stability_d, on = c("variable", "scn", "met")] %>% 
  .[recovery_d, on = c("variable", "scn", "met")] %>% 
  .[, c( "scn", "variable", "description", "unit", "met", "resitance", 
         "resilience", "temp_stability", "recovery")] -> 
  dt

# Add the disturbance severity column. 
dt$`Disturbance Severity` <- paste0(gsub(x = dt$scn,
                                         pattern = "harvest_|_1day_above", 
                                         replacement = ""), " %")

# Format and apply factor information to the variable name so the dimensions of stability plot 
# step through the carbon cycle in order GPP -> NPP -> NEP. 
dt$variable <- gsub(pattern = "_patch", replacement = "", x = dt$variable)
dt$variable <- factor(x = dt$variable, levels = c("GPP", "NPP", "NEP"), ordered = TRUE)
dt <- na.omit(dt)
return(dt)

}

stability_metrics <- calculate_dimensions_stability(annual_data)
```

## Plot Resistance

${r}_{t} = ln(\frac{{F}_{disturbance}}{{F}_{control}})$

When $${t}$$ is the year of the disturbance, in this case 2019.

```{r}
stability_metrics %>% 
  ggplot(aes(`Disturbance Severity`, resitance, color = `Disturbance Severity`)) + 
  geom_point(size = 2, alpha = 0.4) + 
  facet_wrap('variable') + 
  THEME  + 
  labs(title = 'Resistance', y = '2019 ln(F disturabnce / F baseline)') + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  resistance

 ggsave(resistance, filename = file.path(FIG_OUT, "resistance.png"), 
        height = 5,width = 8.5)
 resistance

```

## Plot Resilience

${r}_{i} = ln(\frac{{F}_{disturbance}}{{F}_{control}}) = i + {r}_{i} * t$

```{r}
stability_metrics %>%  
ggplot() + 
  geom_point(aes(`Disturbance Severity`, resilience, color = `Disturbance Severity`),
             size = 2, alpha = 0.4) + 
  facet_wrap('variable') + 
  THEME  + 
  labs(title = 'Resilience', y = 'Slope ln(dist/control)') + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  resilience

 ggsave(resilience, filename = file.path(FIG_OUT, "resilience.png"), 
        height = 5,width = 8.5)
 resilience

```

## Plot Temporal Stability

${s}_{t} = \frac{1}{ SE of resid}$

```{r}
stability_metrics %>%  
ggplot() + 
  geom_point(aes(`Disturbance Severity`, temp_stability, color = `Disturbance Severity`),
             size = 2, alpha = 0.4) + 
  facet_wrap('variable') + 
  THEME  + 
  labs(title = 'Temporal Stability', 
       y = ' 1 / (SE of resid)') + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  temporal_stability

 ggsave(temporal_stability, filename = file.path(FIG_OUT, "temporal_stability.png"), 
        height = 5,width = 8.5)
 temporal_stability
```

## Plot Recovery

Take a look at the recovery at separate time steps.

```{r}
ggplot(data = recovery_at_X_data[ , year := as.character(year)], 
       aes(x = year, y = recovery, fill = `Disturbance Severity`)) + 
  geom_boxplot() + 
  THEME + 
  labs(x = 'Year', y = "ln(disturbance/baseline)", title = "Recovery at Year") + 
  facet_wrap("variable", ncol = 1, scales = "free") + 
  scale_fill_manual(values = FORTE_SEVERITY_COLORS) -> 
  recovery_at
 ggsave(recovery_at, filename = file.path(FIG_OUT, "recovery_at.png"), 
        height = 5,width = 8.5)
 recovery_at

```

Recovery 2 decades after the treatment.

```{r}
stability_metrics %>%  
ggplot() + 
  geom_point(aes(`Disturbance Severity`, recovery, color = `Disturbance Severity`),
             size = 2, alpha = 0.4) + 
  facet_wrap('variable') + 
  THEME  + 
  labs(title = 'Recovery', 
       y = 'Recovery at 2030') + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  recovery_20yrs
 ggsave(recovery_20yrs, filename = file.path(FIG_OUT, "recovery_20yrs.png"), 
        height = 5,width = 8.5)
 recovery_20yrs
```

# Input & Metric Interactions

```{r, message = FALSE, warning=FALSE}
# Create a data frame of the disturbance period annual met inputs, note that it should be 
# a wide data frame in preperation for the pair wise plots. 
data[year == 2019][ , c("variable", "value", "scn")] %>%
  unique() %>% 
  setnames(old = c("scn", "value", "variable"), 
           new = c("met", "met_vavlue", "met_variable")) -> 
  dist_met_data

# Combine the met and the dimensions of stability data into a single data frame. 
stability_metrics[ , c("scn", "variable", "met", "resitance", 
                       "resilience", "temp_stability", "recovery", "Disturbance Severity")] %>%
  melt(id.vars = c("scn", "variable", "met", "Disturbance Severity"),
       measure.vars = c("resitance", "resilience", "temp_stability", "recovery"), 
       variable.name = "metric_name", value.name = "metric_value") %>% 
  unique() -> 
  metric_long

full_df <- dist_met_data[metric_long, on = "met", allow.cartesian=TRUE]
```

```{r}
var_list <- unique(full_df$variable)
metric   <- unique(full_df$metric_name)
full_df  <- full_df[var_info, on = .(met_variable = variable)]

lapply(metric, function(m){
  #  & met_variable %in% c("prate", "tmp", "sh")
  full_df[metric_name == m] %>% 
  ggplot(aes(met_vavlue, metric_value, color = `Disturbance Severity`)) + 
  geom_point() + 
  facet_grid(variable ~ description, scales = "free",  labeller = label_wrap_gen(width = 20, multi_line = TRUE)) + 
  THEME + 
  labs(y = NULL, x = NULL, title = m) + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS)
}) -> 
  plots
names(plots) <- metric
```

## Plot Input x Resitance

```{r, fig.width=6}
plots$resitance + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) -> 
  resistance_inputs

 ggsave(resistance_inputs, filename = file.path(FIG_OUT, "resistance_inputs.png"), width = 20, height = 10)
 resistance_inputs
```

## Plot Input x Resilience

```{r, fig.width=6}
plots$resilience + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
```

## Plot Input x Temporal Stability

```{r, fig.width=6}
plots$temp_stability + 
  labs(title = "Temporal Stability") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
```

## Plot Input x Recovery

```{r, fig.width=6}
plots$recovery + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
```

# Dynamic Recovery

```{r}

# Calculate the ln(disturbance/control)
# Args 
#   d: a data frame of the anual data 
# Return: a data frame of the ln(disturbance/control) per year/variable/met/disturbance severity 
calc_ln_disturbance <- function(d){

  # Check inputs 
  req_names <- c('scn', 'year', 'variable', 'description', 'unit', 'value', 'met')
  assert_that(has_name(d, req_names))

  baseline_scn <- "harvest_0_1day_above" 
  assert_that(baseline_scn %in% d$scn)

  # Separate the control or baseline scneario values from the output from the other 
  # treatment groups. 
  baseline_values <- d[scn == baseline_scn][ , c("year", "variable", "value", "met")]
  names(baseline_values) <- c("year", "variable", "baseline", "met")
  treatment_values <- d[scn != baseline_scn][ , c("scn", "year", "variable", "description", 
                                                "unit", "treatment" = "value", "met")]
  names(treatment_values) <- c("scn", "year", "variable", "description", "unit", "treatment", "met")

  # Combine the control and the treatment values as a wide df, then take the # ln(treatment/control) per year / variable / met
  # realization. 
  disturbance_d <- treatment_values[baseline_values, on = c("year", "variable", "met")][year %in% 2019:2050]
  disturbance_d <- disturbance_d[ , value := log(treatment/baseline)]
  disturbance_d <- disturbance_d[ , disturbance_severity := paste0(gsub(x = scn, pattern = "harvest_|_1day_above", replacement = ""), " %")]
  disturbance_d <- disturbance_d[ , variable :=  gsub(pattern = "_patch", replacement = "", x = variable)]
  disturbance_d$variable <- factor(x = disturbance_d$variable, levels = c("GPP", "NPP", "NEP"), ordered = TRUE)
  disturbance_d <- disturbance_d[ , .(year, variable, met, value, disturbance_severity)]
  disturbance_d <- na.omit(disturbance_d)
  disturbance_d <- disturbance_d[year %in% 2019:2050]

  return(disturbance_d)
}

disturbance <- calc_ln_disturbance(annual_data)

# Define the functions to be used by frollapply to get the fits from the rolling slopes 
slope_func <- function(value){

  assert_that(is.vector(value))
  sq <- seq(1:length(value))
  fit <- lm(value~sq)

  slope <- coef(fit)[2]
  names(slope) <- "slope"
  return(slope)

}
intercept_func <- function(value){

  assert_that(is.vector(value))
  sq <- seq(1:length(value))
  fit <- lm(value~sq)

  inter <- coef(fit)[1]
  names(inter) <- "y-intercept"
  return(inter)

}

# Apply the slope and intercept functions to each variable / met / severity disturbance values. 
split(disturbance,
      interaction(disturbance$variable, disturbance$met, disturbance$disturbance_severity)) %>% 
  lapply(function(x){

    window <- 5
    slope <- frollapply(x = x[["value"]], n = window, FUN = slope_func, align = "left")
    slope2 <- frollapply(x = slope, n = window, FUN = slope_func, align = "left")

    out <- cbind(x, slope = slope, slope2 = slope2)
    return(out)

  }) %>% 
  rbindlist %>%  
  na.omit -> 
  dist_slope_y
```

## Plot of Disturbance Rate of Change

This is the first derivative, the rate of change of disturabnce, we were intrested in when this is 0, if that would be the recovery point becacuse that would refelect when the ecosystem has reached a new steady state...

```{r}
ggplot(data = dist_slope_y[year %in% 2019:2050]) +
 geom_hline(aes(yintercept = 0)) +
 geom_line(aes(year, slope, group = interaction(disturbance_severity, met)), color = "grey", alpha = 0.3) +
 geom_point(aes(year, slope, color = disturbance_severity, group = interaction(disturbance_severity, met))) +
 facet_wrap("variable", scales = "free", ncol = 1) +
 theme_bw() +
 labs(title = "5 year running slope")
```

```{r}
# Selet the recovery years as the years where the mangnitude of the rate of change is less than 1e-3 or set a 
# hard limit of 2030, which is unideal but for now it is not not to avoid some sort of arbitriaty cut off. 
recovery_yr <- dist_slope_y[abs(slope) <= 1e-3][ , .(recovery_yr = min(year)), by = .(variable, met, disturbance_severity)] 
recovery_yr <- recovery_yr[recovery_yr <= 2030]

# Now join with the disurbance values. 
recovery_yr[disturbance, on = c("variable", "met", "disturbance_severity"), nomatch = NA] %>% 
  .[ , recovery_yr:= ifelse(test = is.na(recovery_yr), yes = 2030, no = recovery_yr)] %>% 
  .[year <= recovery_yr] -> 
  disturbance_recovery_subset

# Find recovery value aka the disturbance value at the time of recovery. 
recovery_values <- disturbance_recovery_subset[ year == recovery_yr]
```

Using when the rate of change or running slope is less than or equal to 0 or with a cut off there are only a handful of runs that recovery or reach a new stead state.

```{r}
recovery_yr[disturbance, on = c("variable", "met", "disturbance_severity"), nomatch = NA] %>% 
 # .[ , recovery_yr:= ifelse(test = is.na(recovery_yr), yes = 2030, no = recovery_yr)] %>% 
  .[year <= recovery_yr] %>% 
  ggplot(aes(year, value, color = disturbance_severity, group = interaction(met, disturbance_severity))) + 
  geom_line() + 
  facet_grid(variable~disturbance_severity, scales = "free") + 
  labs(title = "Dynamic Recovery", y = "ln(treatment/control)") + 
  theme_bw() + 
  theme(legend.position = "none")
```

What if the cut off value is when the rate off change is less than or equal to 0 or at most 2030? If that is the case then our results look something like.

```{r}
recovery_yr[disturbance, on = c("variable", "met", "disturbance_severity"), nomatch = NA] %>%
  .[ , recovery_yr:= ifelse(test = is.na(recovery_yr), yes = 2030, no = recovery_yr)] %>% 
  .[year <= recovery_yr] %>% 
  ggplot(aes(year, value, color = disturbance_severity, group = interaction(met, disturbance_severity))) + 
  geom_line() + 
  facet_grid(variable~disturbance_severity, scales = "free") + 
  labs(title = "Dynamic Recovery", y = "ln(treatment/control)") + 
  theme_bw() + 
  theme(legend.position = "none")
```

Is a linear fit the right fit here? How do we deal with these different recovery dates?

### 2021-03-03 C.analysis/plan.R

@@ -1,300 +1,57 @@
-# Define drake plans that are used to process/format the ED experiment results. 
-
-# 0. Set Up ------------------------------------------------------------------
 source(here::here('C.analysis', 'packages.R'))
-source(here::here('C.analysis', 'functions.R'))
+source(here::here('C.analysis', 'ED_functions.R'))
 
 ED_OUTPUT_DIR <- here::here('ED-outputs')
+WRITE_TO      <-  here::here('ED-outputs', 'results')
+dir.create(WRITE_TO, showWarnings = FALSE)
 
-# The helper functions that chain together the 
-prep_data <- function(path, scn){
-  
-  assertthat::assert_that(file.exists(path))
-  object <- readRDS(path)
-  
-  data <- list(NPP = process_NPP(object, scn), 
-               LAI = process_LAI(object, scn), 
-               NEP = process_NEP(object, scn), 
-               ABG = process_ABG(object, scn), 
-               GPP = process_GPP(object, scn))
-  
-  return(data)
-}
-format_list <- function(l){
-  
-  assertthat::assert_that(is.list(l))
-  list('NPP' = ed_format_list(name = 'NPP', to_merge = l),
-       'LAI' = ed_format_list(name = 'LAI', to_merge = l), 
-       'NEP' = ed_format_list(name = 'NEP', to_merge = l), 
-       'ABG' = ed_format_list(name = 'ABG', to_merge = l), 
-       'GPP' = ed_format_list(name = 'GPP', to_merge = l))
-}
+#  This is the old experimentn 1 (prior to the met ensemble) work flow
+# exp1_plan <- drake_plan(exp = "exp-1", 
+#            files = list.files(file.path(ED_OUTPUT_DIR, exp), pattern = '.rds', full.names = TRUE), 
+#            scns = get_scn_names(files),
+#            
+#            # Process the monthly values 
+#            monthly = rbindlist(mapply(get_monthly_data, path = files, scn = scns, SIMPLIFY = FALSE)),
+#            mon_data = target({
+#              fname = file.path(WRITE_TO, paste0(exp, '-mon.csv'))
+#              write.csv(monthly, file = fname, row.names = FALSE)
+#              }), 
+#            
+#            # Calculate the annual values 
+#            annual = calculate_annual_values(monthly), 
+#            an_data = target({
[…]

### 2021-03-11 C.analysis/exp-1_met.Rmd
Materials and code for the March VCU meeting

@@
 library(knitr)
+library(kableExtra)
 
@@
 
+Create a summary table of the disturbance, the inital response to the disturbance
+
 ```{r}
+
+annual_data[year %in% c(2018, 2019), ] %>% 
+  .[scn != "harvest_0_1day_above"] %>%  
+  .[, .(scn, year, variable, value, met)] %>% 
+  dcast(scn + variable + met ~ year) %>%  
+  .[ , percent_change := 100 * (`2019` - `2018`) / `2018`] %>% 
+  # summary stats!! 
+  .[ , .(min = min(percent_change), 
+         max = max(percent_change),
+         mean = mean(percent_change), 
+         sd = sd(percent_change)), by = .(scn, variable)] %>%  
+  .[ , .(variable, scn, min, max, mean, sd)] -> 
+summary_data
+
+summary_data$variable <- gsub(pattern = "_patch",
+                              replacement  = "",
+                              x = summary_data$variable)
+summary_data$severity <- gsub(pattern = "harvest_|_1day_above",
+                              replacement = "", 
+                              x = summary_data$scn)
+summary_data[ , .(variable, severity, min, max, mean, sd)] %>%  
+  setorder(variable) %>% 
+  knitr::kable(format = "html", digits = 2) %>% 
+  collapse_rows(columns = 1, valign = "top")
+
+```
+
+```{r}
+annual_data[variable == "GPP_patch", ] %>% 
+  .[year %in% c(2018:2021), ] %>% 
+  ggplot(aes(year, value, color = scn)) + 
+  geom_point()
+
+```
+
+```{r}
+# Plot the individual disturbance severities
+
 # dt$`Disturbance Severity` <- paste0(dt$severity, ' %')
@@
 
+  ggsave(recovery_at, filename = file.path(FIG_OUT, "recovery_at.png"), 
+        height = 5,width = 8.5)
+ recovery_at
+
 ```
@@
   full_df[metric_name == m] %>% 
+    .[met_variable %in%c("prate", "tmp", "sh")]%>% 
   ggplot(aes(met_vavlue, metric_value, color = `Disturbance Severity`)) + 
@@
 
- ggsave(resistance_inputs, filename = file.path(FIG_OUT, "resistance_inputs.png"), width = 20, height = 10)
+ ggsave(resistance_inputs, filename = file.path(FIG_OUT, "resistance_inputs.png"), width = 5, height = 5)
  resistance_inputs

### 2021-03-11 C.analysis/plan.R

@@ -5,28 +5,7 @@ ED_OUTPUT_DIR <- here::here('ED-outputs')
 WRITE_TO      <-  here::here('ED-outputs', 'results')
 dir.create(WRITE_TO, showWarnings = FALSE)
 
-#  This is the old experimentn 1 (prior to the met ensemble) work flow
-# exp1_plan <- drake_plan(exp = "exp-1", 
-#            files = list.files(file.path(ED_OUTPUT_DIR, exp), pattern = '.rds', full.names = TRUE), 
-#            scns = get_scn_names(files),
-#            
-#            # Process the monthly values 
-#            monthly = rbindlist(mapply(get_monthly_data, path = files, scn = scns, SIMPLIFY = FALSE)),
-#            mon_data = target({
-#              fname = file.path(WRITE_TO, paste0(exp, '-mon.csv'))
-#              write.csv(monthly, file = fname, row.names = FALSE)
-#              }), 
-#            
-#            # Calculate the annual values 
-#            annual = calculate_annual_values(monthly), 
-#            an_data = target({
-#              fname = file.path(WRITE_TO, paste0(exp, '-yr.csv'))
-#              write.csv(annual, file = fname, row.names = FALSE)
-#            })
-# ) 
-#            
-       
-
+         
 # Thoughts about this work flow, why doesn't drake trigger when more files were added? 
 exp1_met_plan <- drake_plan(exp = "exp-1", 
                         files = list.files(file.path(ED_OUTPUT_DIR, exp), pattern = '.rds', full.names = TRUE), 

### 2021-03-11 make.R

@@ -8,4 +8,4 @@ source("C.analysis/ED_functions.R") # Define your custom code as a bunch of func
 source("C.analysis/plan.R")      # Create your drake plan.
 
 
-make(exp1_plan)
+make(exp1_met_plan)

### 2021-05-06 A.inputs/1.generate_met_ensemble.R
clean up

@@ -1,99 +1,0 @@
-# Note that this requires that 1A.met_input_processing.R has already been run.
-
-# Load the libraries 
-library(magrittr)
-library(here)
-library(data.table)
-library(ncdf4)
-library(purrr)
-library(lubridate)
-library(assertthat)
-library(hdf5r)
-
-# There were some issues installing pecan specfic scripts should be sourced from the pecan projects. 
-# See https://github.com/PecanProject/pecan to downlaod peacn 
-# git show-branch 
-# [develop] Merge pull request #2570 from infotroph/book-headers
-# commit b264ce700a1892726a7e4ec21c3187152297a190 
-# brach devlop
-# These files have been committed to this repo for purposes of reproducibility. 
-# TODO is this legal or no? 
-BASE_DIR  <- here::here()
-#PECAN_DIR <- '/people/dorh012/forte-workflow/pecan'
-#assertthat::assert_that(dir.exists(PECAN_DIR))
-#source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'read_ed2in.R'))
-#source(file.path(PECAN_DIR,  'models', 'ed', 'R', 'write_ed2in.R'))
-
-# Define the directories
-BASE_DIR <- here::here()
-MET_DIR  <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
-WRITE_TO <- here::here("ED-outputs", "met_results")
-
-# Make sure that the met data dir exists. 
-assertthat::assert_that(dir.exists(MET_DIR))
-
-source(here::here('A.inputs', 'helper_functions.R'))
-
-# Prepare a met directory 
-# Args
-#   out_dir: str path of the directory 
-# Return: NULL, copy over a set of files 
-prep_dir <- function(out_dir){
-  
-  assert_that(dir.exists(out_dir))
-  
-  modify_met_header(out_dir)
-  
-  files_to_copy <- get_year_month(list.files(file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main'),
-                                             pattern = 'h5', full.names = TRUE))
-  more_files_to_copy <- get_year_month(list.files(file.path(BASE_DIR, 'A.inputs', 'NARR-ED2'), 
-                                                  pattern = 'h5', full.names = TRUE))[year <= min(files_to_copy$year)]
-  
-  files_to_copy <- c(files_to_copy$name, more_files_to_copy$name)
[…]

### 2021-05-06 B.ed/1.Experiment-1.R

@@ -52,6 +52,10 @@ case <- merge(case, met_headers)
 case$casename <- paste(case$casename, case$met_name, sep = '_')
 case <- case[c("casename", "IYEARA", "IYEARZ", "EVENT_FILE", "ED_MET_DRIVER_DB")]
 
+case <- case[!grepl(x = case$casename, pattern = "harvest_0_1day_above"), ]
+
+
+
 # For each entry in the case data frame generate the ed run set up. 
 for(i in 1:nrow(case)){ 
   setup_ed_run(case[i,], write_to = WRITE_TO) 

### 2021-05-06 B.ed/2.post_processing.R

@@ -15,10 +15,10 @@ source(to_source)
 
 # 1. Process output directories -----------------------------------------------------------------
 # Start with experiment 1! And move them into the output direcotry 
-files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'exp-1')
-exp1_out <- file.path(OUT_DIR, 'exp-1')
+files    <- process_outputs(BASE_DIR = BASE_DIR, exp_dirs = 'exp-constant')
+exp1_out <- file.path(OUT_DIR, 'exp-constant')
 dir.create(exp1_out)
-rds_files <- list.files(file.path(BASE_DIR, 'exp-1'), pattern = '.rds', recursive = TRUE, full.names = TRUE)
+rds_files <- list.files(file.path(BASE_DIR, 'exp-constant'), pattern = '.rds', recursive = TRUE, full.names = TRUE)
 system2('cp', args = c(paste0(rds_files, collapse = ' '), exp1_out))
 
 message('move the ED outputs with scp')

### 2021-05-06 C.analysis/ED_functions.R

@@ -1,361 +1,0 @@
-# Custom functions for the drake workflow that is used to process the 
-# monthly ED output files into data frames that are easy to work with. 
-
-
-# Add the ED pft names of the data table 
-#   Args  
-#     dt: the data table of ED information, it must have PFT column 
-#   Return: a data table with the pft names 
-add_pft_fullnames <- function(dt){
-  
-  # Check the inputs 
-  assert_that(is.data.table(dt))
-  assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
-  
-  # Data table of the plant functional types in ED. 
-  pft_names <- data.table(pft = c(6, 8, 9, 10, 11, 7) , 
-                          pft_name = c('Northern North American pines', 
-                                       'Late-successional North American conifers', 
-                                       'Temperate broadleaf, early successional',  
-                                       'Temperate broadleaf, mid-successional', 
-                                       'Temperate broadleaf, late successional', 
-                                       'Southern North American pines'))
-  
-  
-  # Add the pft names to the pft data table. 
-  named_dt <- na.omit(dt[pft_names, on = c("PFT" = "pft")])
-  return(named_dt)
-  
-}
-
-
-# Add FoRTE variable information to the data table 
-#   Args
-#     dt: a data table of ED output 
-# Return: a data table of ED output with the the full variable name and unit information
-add_var_info <- function(dt){
-  
-  # Check inputs 
-  assert_that(is.data.table(dt))
-  assert_that('variable' %in% names(dt), msg = 'dt is missing the column named variable')
-  assert_that(unique(dt[['variable']]) %in% ed4forte::ed2_variable_info()[['variable']], msg = 'dt variable is missing from ed4forte::ed2_variable_info')
-  
-  # Add the variable information to the data table. 
-  dt <- dt[as.data.table(ed4forte::ed2_variable_info())[, list(variable, description, unit)], on = 'variable',  nomatch=0]
-  return(dt)
-  
-}
-
-
-# Label the cohorts 
-# Args 
-#   dt: the data table of the cohort data 
-# Return: a data frame of the data with cohort information 
-add_cohort_labels <- function(dt){
[…]

### 2021-05-06 C.analysis/packages.R

@@ -1,9 +1,0 @@
-# Load all your packages before calling make().
-
-library(drake)
-library(tibble)
-library(data.table)
-library(lubridate)
-library(assertthat)
-library(ed4forte)
-

### 2021-05-06 C.analysis/plan.R

@@ -1,36 +1,70 @@
-source(here::here('C.analysis', 'packages.R'))
-source(here::here('C.analysis', 'ED_functions.R'))
+# Load helper functions. 
+source(here::here('C.analysis', '0.packages.R'))
+source(here::here('C.analysis', '0.func_ED.R'))
+source(here::here('C.analysis', '0.func_metric.R'))
 
+
+# Define the inputs and output files. 
 ED_OUTPUT_DIR <- here::here('ED-outputs')
 WRITE_TO      <-  here::here('ED-outputs', 'results')
 dir.create(WRITE_TO, showWarnings = FALSE)
 
-         
-# Thoughts about this work flow, why doesn't drake trigger when more files were added? 
-exp1_met_plan <- drake_plan(exp = "exp-1", 
-                        files = list.files(file.path(ED_OUTPUT_DIR, exp), pattern = '.rds', full.names = TRUE), 
-                        scns = get_scn_names(files),
-                        
-                        # Process the monthly values 
-                        monthly = rbindlist(mapply(get_monthly_data, path = files, scn = scns, SIMPLIFY = FALSE)),
-                        mon_data = target({
-                          # TODO 
-                          # Parse out the met name, this might want to go into a function
-                          monthly <- monthly[, met := stringr::str_sub(scn, -4)]
-                          monthly <- monthly[, scn := gsub('.{5}$', '', scn)]
-                          
-                          fname = file.path(WRITE_TO, paste0(exp, '-mon.csv'))
-                          write.csv(monthly, file = fname, row.names = FALSE)
-                        }), 
-                        
-                        # Calculate the annual values 
-                        annual = calculate_annual_values(monthly), 
-                        an_data = target({
-                          # TODO this might need to be made innto a function 
-                          # Parse out the met name, this might want to go into a function
-                          annual <- annual[, met := stringr::str_sub(scn, -4)]
-                          annual <- annual[, scn := gsub('.{5}$', '', scn)]
-                          fname = file.path(WRITE_TO, paste0(exp, '-yr.csv'))
-                          write.csv(annual, file = fname, row.names = FALSE)
-                        })
-)     
+# Process ED output into csv files ---------------------------------------------------------------------
[…]

### 2021-05-06 A.inputs/1A.constant_met_inputs.R
update the A.inputs

@@ -1,0 +1,121 @@
+# Looking at the ED results in the baseline scenario it is somewhat concerning
+# that we see this periodicity in the historical period, is that a function of ED
+# or is that a function of the meterolgical inputs. Another outstanding question is 
+# why the climate ensemble begins to converge. It is a coding problem or a numerical 
+# feature of ED? 
+
+# 0. Set UP --------------------------------------------------------------------------
+# Libraries 
+library(magrittr)
+library(here)
+library(data.table)
+library(ncdf4)
+library(purrr)
+library(lubridate)
+library(assertthat)
+library(hdf5r)
+library(assertthat)
+
+
+# Define the directories
+BASE_DIR <- here::here()
+MET_DIR  <- file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main')
+WRITE_TO <- here::here("ED-outputs", "met_results")
+
+# Make sure that the met data dir exists. 
+assertthat::assert_that(dir.exists(MET_DIR))
+source(here::here('A.inputs', 'helper_functions.R'))
+
+# Prepare a met directory 
+# Args
+#   out_dir: str path of the directory 
+# Return: NULL, copy over a set of files 
+prep_dir <- function(out_dir){
+  
+  assert_that(dir.exists(out_dir))
+  
+  modify_met_header(out_dir)
+  
+  files_to_copy <- get_year_month(list.files(file.path(BASE_DIR, 'A.inputs', 'NARR-ED2-main'),
+                                             pattern = 'h5', full.names = TRUE))
+  more_files_to_copy <- get_year_month(list.files(file.path(BASE_DIR, 'A.inputs', 'NARR-ED2'), 
+                                                  pattern = 'h5', full.names = TRUE))[year <= min(files_to_copy$year)]
+  
+  files_to_copy <- c(files_to_copy$name, more_files_to_copy$name)
+  out <- copy_files_func(files_to_copy, out_dir)
+  
+}
+
+
+# 1. Set up the input files -------------------------------------------------------------------
+# Set up the years of data to used 
+years <- floor(seq(from = 1979, to = 2019, length.out = 20)) 
+tag       <- paste0("_met-", years)
+dir_names <- paste0("NARR-ED2", tag)
+
+# Come up with the names of the met dir. 
+# met_dirs  <- list.files(path = file.path(BASE_DIR, 'A.inputs'), pattern = paste0(dir_names, collapse = "|"), full.names = TRUE)
[…]

### 2021-05-06 C.analysis/paper_analysis.Rmd
add paper1 materials

---
title: "FoRTE Disturbance Paper Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Objective

## Set up script

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)

library(dplyr)
library(ggplot2)

BASE_DIR <- "/Users/dorh012/Documents/2020/FoRTE/forte-disturbance"
INPUT_DIR <- file.path(BASE_DIR, "ED-outputs", "results")

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
FORTE_SEVERITY_COLORS <- c("#000000", "#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('0 %', '45 %', '65 %', '85 %')
```

Load the data

```{r}
# Load the ED output data
ED_output <- read.csv(list.files(INPUT_DIR, "exp-constant-yr.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the meteorological input, the annual average and the monthly average. 
annual_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data.csv", full.names = TRUE), stringsAsFactors = FALSE)
monthly_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data-monthly.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the results realted to the metrics of ecosystem stability, the log ratio of the time series, resilience, 
# and resistance. 
disturbance <- read.csv(file.path(INPUT_DIR, "metric", "ln_ratio_timeseries.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resilience <- read.csv(file.path(INPUT_DIR, "metric", "resilience_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resistance <- read.csv(file.path(INPUT_DIR, "metric", "resistance_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
```

# Baseline Scenario

```{r}
ED_output %>%  
  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
  mutate(description = paste0(variable, ' (', unit,  ')')) %>% 
  ggplot(aes(year, value, color = met, group = met)) + 
  geom_line() + 
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
  THEME +
  labs(y = "")
```

# Disturbance Time Series

Log ratio plots show in the Hillbrand manuscript.

```{r}
ggplot(data = disturbance) + 
  geom_line(aes(year, value, color = severity, line = met), alpha = 0.5) + 
  facet_wrap("variable", scales = 'free') +
  THEME +
  labs(y = "ln(treatment/control)") +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  labs(title = "Disturbance Response")
```

# Resistance

```{r}
ggplot(data = resistance) + 
  geom_point(aes(severity, trough_resistance, color = severity), position = position_jitter(), alpha = 0.5, size = 2) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "ln(dist/control) unitless", 
       title = "Resistance") + 
  THEME + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS)

```

# Resilience

```{r}
ggplot(data = resilience) + 
  geom_point(aes(severity, resilience, color = severity), position = position_jitter(), alpha = 0.5, size = 2) + 
  facet_wrap("variable", scales = "free") + 
  labs(y = "ln(dist/control) unitless", 
       title = "Resilience") + 
  THEME + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS)
```

# Resistance vs Resilience

```{r}
resilience %>%
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.5, size = 2) + 
   facet_wrap("variable", scales = "free") + 
  labs(x = "resistance", 
       y = "resilience", 
       title = "Resilience vs. Resistance") + 
  THEME + 
  scale_color_manual(values = FORTE_SEVERITY_COLORS)
```

# Investigating the relationship between met and the metrics

Annual met values (by design they should be constant/flat over the entire time period)

```{r, fig.width=8, fig.height=8}
annual_met %>% 
  mutate(description = paste0(description, ' (', units, ')')) %>% 
  ggplot(aes(year, value, color = met, group = met)) + 
  geom_line() +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
  THEME +
  labs(y = NULL, x  = NULL,  title = "Annual Met") + 
  theme(legend.position = "none")
```

Monthly values, take a look at the pattern that is going on within each year.

```{r, fig.width=8, fig.height=8}
monthly_met %>% 
  mutate(description = paste0(description, ' (', units, ')')) %>% 
  ggplot(aes(mon_time, value, color = met, group = met, line = met)) + 
  geom_line() +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
  THEME +
  labs(y = NULL, x  = NULL, title = "Monthly Met") + 
  theme(legend.position = "none")
```

```{r}
annual_met %>%  
  mutate(met = as.integer(met)) %>% 
  select(met, met_var = variable, met_value = value, met_des = description) %>% 
  distinct() -> 
  met_data

resilience %>%
  left_join(resistance) %>%  
  mutate(met = as.integer(met)) %>% 
  left_join(met_data) %>%  
  select(scn, severity, met, variable, resilience, resistance  = trough_resistance, met_var, met_value, met_des) -> 
  metrics_v_met
```

## Met vs. Metrics  {.tabset}

```{r}
# Save a copy of the meteorological variables 
vars <- unique(metrics_v_met$met_var)

# Create the plots of met vs resistance. 
lapply(vars, function(met_var_name){

  des <- unique(metrics_v_met[metrics_v_met$met_var == met_var_name, ]$met_des)

  metrics_v_met %>% 
    filter(met_var == met_var_name) %>%  
    ggplot(aes(met_value, resistance, color = severity)) + 
    geom_point() + 
    facet_wrap('variable', scales = "free", nrow = 1) + 
    THEME + 
    scale_color_manual(values = FORTE_SEVERITY_COLORS) +
    labs(x  = des, title = 'Resistance') -> 
    out

  return(out)

}) ->
  resistance_plots
names(resistance_plots) <- vars

# Create the plots of met vs resilience 
lapply(vars, function(met_var_name){

  des <- unique(metrics_v_met[metrics_v_met$met_var == met_var_name, ]$met_des)

  metrics_v_met %>% 
    filter(met_var == met_var_name) %>%  
    ggplot(aes(met_value, resilience, color = severity)) + 
    geom_point() + 
    facet_wrap('variable', scales = "free", nrow = 1) + 
    THEME + 
    scale_color_manual(values = FORTE_SEVERITY_COLORS) +
    labs(x  = des, title = 'Resilience') -> 
    out

  return(out)

}) ->
  resilience_plots
names(resilience_plots) <- vars
```

### downward long wave radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$dlwrf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$dlwrf
```

### near infrared beam downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$nbdsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$nbdsf
```

### near IR diffuse downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$nddsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$nddsf
```

### visible beam downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$vbdsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vbdsf
```

### visible diffuse downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$vddsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vddsf
```

### precipitation rate

```{r, fig.width=10, fig.height=4}
resistance_plots$prate
```

```{r, fig.width=10, fig.height=4}
resilience_plots$prate
```

### atmospheric pressure

```{r, fig.width=10, fig.height=4}
resistance_plots$pres
```

```{r, fig.width=10, fig.height=4}
resilience_plots$pres
```

### geopotential height

```{r, fig.width=10, fig.height=4}
resistance_plots$hgt
```

```{r, fig.width=10, fig.height=4}
resilience_plots$hgt
```

### zonal wind

```{r, fig.width=10, fig.height=4}
resistance_plots$ugrd
```

```{r, fig.width=10, fig.height=4}
resilience_plots$ugrd
```

### meridional wind

```{r, fig.width=10, fig.height=4}
resistance_plots$vgrd
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vgrd
```

### specific humidity

```{r, fig.width=10, fig.height=4}
resistance_plots$sh
```

```{r, fig.width=10, fig.height=4}
resilience_plots$sh
```

### air temperature

```{r, fig.width=10, fig.height=4}
resistance_plots$tmp
```

```{r, fig.width=10, fig.height=4}
resilience_plots$tmp
```

## Met vs Metrics observations

Breifly looking through the different pannels what I've notcies is that patterns between meterology values and the stability meteric are more linear looking for resistance comapred to resillience.

The three meterology variables that are the most interesting visually are

* air temperature 
* precipitation rate
* air temperature

# TO DO

Kalyn do you want to take a stab that a random variancne

### 2021-05-06 B.ed/1.Experiment-1.R

@@ -1,67 +1,0 @@
-## FoRTE ED Disturbence: Experiment  1 
-## Bare-ground, 100 year run to mature stand, 0-45-65-85 harvest of C storage pool, 
-## run for 10 more years, using climatology (average of 5 years of data), ED default parameters
-## monthly NPP, NEE, LAI, biomass by cohort later we'll want to run multiple times to average out randomness
-## See https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77
-
-# 0. Set Up  ------------------------------------------------------------------------------------
-library(data.table)
-
-# Define directories. 
-BASE_DIR  <- here::here()
-WRITE_TO  <- file.path(BASE_DIR, 'exp-1'); dir.create(WRITE_TO, showWarnings = FALSE, recursive = TRUE) # This is the directory to write where the different cases should be set up. 
-EVENT_DIR <- file.path(BASE_DIR, 'A.inputs', 'events')
-
-# Load the required pacakges and functions. 
-to_source <- file.path(BASE_DIR, "B.ed", "0A.dependencies_functions.R")
-stopifnot(file.exists(to_source))
-source(to_source)
-
-
-# 1. Define ED Runs ------------------------------------------------------------------------------------
-# Use the case data frame to set up the different runs with the 0-45-65-85 harvest of the C storage pool. 
-
-# Start by defining the ED run information. 
-IYEARA <- 1900
-IYEARZ <- 2049
-
-# Find the event files
-event_file <- list.files(EVENT_DIR, pattern = '1day_above.xml', full.names = TRUE)
-cases      <- gsub(pattern = '.xml', replacement = '', x = basename(event_file))
-assertthat::assert_that(all(file.exists(event_file)))
-# Important! When reading in the event_file paths they must be written out as a string, 
-# otherwise ED will fail but will throw an uninformative error message realted to 
-# missing a different ED input. 
-event_file <- paste0("'", event_file, "'") 
-
-# Save all of the ED information as a single data frame. 
-case <- data.frame(casename = cases, IYEARA = IYEARA, IYEARZ = IYEARZ, EVENT_FILE = event_file)
-
-
-met_headers <- data.table(ED_MET_DRIVER_DB = c(file.path(INPUT_DIR, "NARR-ED2_met1", "ED_MET_DRIVER_HEADER"), 
[…]

### 2021-05-06 B.ed/1.constant-met-runs.R

@@ -1,0 +1,63 @@
+## FoRTE ED Disturbence: Experiment  1 
+## Bare-ground, 100 year run to mature stand, 0-45-65-85 harvest of C storage pool, 
+## run for 10 more years, using climatology (average of 5 years of data), ED default parameters
+## monthly NPP, NEE, LAI, biomass by cohort later we'll want to run multiple times to average out randomness
+## See https://github.com/FoRTExperiment/FoRTE-mgmt/issues/77
+
+# 0. Set Up  ------------------------------------------------------------------------------------
+library(data.table)
+
+# Define directories. 
+BASE_DIR  <- here::here()
+WRITE_TO  <- file.path(BASE_DIR, 'exp-constant'); dir.create(WRITE_TO, showWarnings = FALSE, recursive = TRUE) # This is the directory to write where the different cases should be set up. 
+EVENT_DIR <- file.path(BASE_DIR, 'A.inputs', 'events')
+
+# Load the required pacakges and functions. 
+to_source <- file.path(BASE_DIR, "B.ed", "0A.dependencies_functions.R")
+stopifnot(file.exists(to_source))
+source(to_source)
+
+
+# 1. Define ED Runs ------------------------------------------------------------------------------------
+# Use the case data frame to set up the different runs with the 0-45-65-85 harvest of the C storage pool. 
+
+# Start by defining the ED run information. 
+IYEARA <- 1900
+IYEARZ <- 2050
+
+# Find the event files
+event_file <- list.files(EVENT_DIR, pattern = '1day_above.xml', full.names = TRUE)
+cases      <- gsub(pattern = '.xml', replacement = '', x = basename(event_file))
+assertthat::assert_that(all(file.exists(event_file)))
+# Important! When reading in the event_file paths they must be written out as a string, 
+# otherwise ED will fail but will throw an uninformative error message realted to 
+# missing a different ED input. 
+event_file <- paste0("'", event_file, "'") 
+
+# Save all of the ED information as a single data frame. 
+case <- data.frame(casename = cases, IYEARA = IYEARA, IYEARZ = IYEARZ, EVENT_FILE = event_file)
+
+met_dirs <- list.files(INPUT_DIR, "constant", full.names = TRUE)
+met_headers <- file.path(met_dirs, "ED_MET_DRIVER_HEADER")
[…]

### 2021-05-06 B.ed/1.constant_pulse_met.R

@@ -1,0 +1,48 @@
+## FoRTE ED Disturbence: met ensemble
+## Bare-ground, 100 year run to mature stand constant met with the expection of a different met pulse 
+## duirng a 1, 2, 5, and 10 year periods. 
+## question is it wrong that during the "pulse" the met is changing and is not held constant. 
+## Hmmm what am I trying to answer with this? I guess I am intrested in how long the met has a memory/legacy 
+## effect. Let's jsut start with the 1 yr and 5 year run. 
+
+library(data.table)
+
+# Define directories. 
+BASE_DIR  <- here::here()
+WRITE_TO <- file.path(BASE_DIR, "constant-pulse-met")
+dir.create(WRITE_TO)
+  
+# Load the required pacakges and functions. 
+to_source <- file.path(BASE_DIR, "B.ed", "0A.dependencies_functions.R")
+stopifnot(file.exists(to_source))
+source(to_source)
+
+
+
+# 1. Define ED Runs ------------------------------------------------------------------------------------
+
+# Start by defining the ED run information. 
+IYEARA <- 1900
+IYEARZ <- 2020
+
+
+# Save all of the ED information as a single data frame. 
+case <- data.frame(IYEARA = IYEARA, IYEARZ = IYEARZ)
+met_headers <- data.table(ED_MET_DRIVER_DB = c(file.path(INPUT_DIR, "NARR-ED2met1-constant_1yr", "ED_MET_DRIVER_HEADER"), 
+                                               file.path(INPUT_DIR, "NARR-ED2met1-constant_5yr", "ED_MET_DRIVER_HEADER")), 
+              met_name = c('1_yr', '5_yr'))
+
+case <- merge(case, met_headers)
+case$casename <- case$met_name
+as.data.frame(case) %>% 
+  dplyr::select("casename", "IYEARA", "IYEARZ", "ED_MET_DRIVER_DB") %>%  
+  as.data.table -> 
+  case
+
+# For each entry in the case data frame generate the ed run set up. 
+for(i in 1:nrow(case)){ 
+  setup_ed_run(case[i,], write_to = WRITE_TO) 
+}
+
+# Create the sh script. 
+setup_sh(WRITE_TO)

### 2021-05-06 C.analysis/0.func_ED.R

@@ -1,0 +1,440 @@
+# Custom functions for the drake workflow that is used to process the 
+# monthly ED output files into data frames that are easy to work with. 
+
+
+# Add the ED pft names of the data table 
+#   Args  
+#     dt: the data table of ED information, it must have PFT column 
+#   Return: a data table with the pft names 
+add_pft_fullnames <- function(dt){
+  
+  # Check the inputs 
+  assert_that(is.data.table(dt))
+  assert_that('PFT' %in% names(dt), msg = 'PFT column is missing from the dt')
+  
+  # Data table of the plant functional types in ED. 
+  pft_names <- data.table(pft = c(6, 8, 9, 10, 11, 7) , 
+                          pft_name = c('Northern North American pines', 
+                                       'Late-successional North American conifers', 
+                                       'Temperate broadleaf, early successional',  
+                                       'Temperate broadleaf, mid-successional', 
+                                       'Temperate broadleaf, late successional', 
+                                       'Southern North American pines'))
+  
+  
+  # Add the pft names to the pft data table. 
+  named_dt <- na.omit(dt[pft_names, on = c("PFT" = "pft")])
+  return(named_dt)
+  
+}
+
+
+# Add FoRTE variable information to the data table 
+#   Args
+#     dt: a data table of ED output 
+# Return: a data table of ED output with the the full variable name and unit information
+add_var_info <- function(dt){
+  
+  # Check inputs 
+  assert_that(is.data.table(dt))
+  assert_that('variable' %in% names(dt), msg = 'dt is missing the column named variable')
+  assert_that(unique(dt[['variable']]) %in% ed4forte::ed2_variable_info()[['variable']], msg = 'dt variable is missing from ed4forte::ed2_variable_info')
+  
+  # Add the variable information to the data table. 
+  dt <- dt[as.data.table(ed4forte::ed2_variable_info())[, list(variable, description, unit)], on = 'variable',  nomatch=0]
+  return(dt)
+  
+}
+
+
+# Label the cohorts 
+# Args 
+#   dt: the data table of the cohort data 
+# Return: a data frame of the data with cohort information 
[…]

### 2021-05-06 C.analysis/0.packages.R

@@ -1,0 +1,10 @@
+# Load all your packages before calling make().
+
+library(drake)
+library(tibble)
+library(data.table)
+library(lubridate)
+library(assertthat)
+library(ed4forte)
+library(segmented)
+library(dplyr)

### 2021-05-06 make.R

@@ -3,9 +3,20 @@
 # It is not the only way to organize R scripts for drake,
 # but this pattern is helpful.
 
-source("C.analysis/packages.R")  # Load your packages, e.g. library(drake).
-source("C.analysis/ED_functions.R") # Define your custom code as a bunch of functions.
-source("C.analysis/plan.R")      # Create your drake plan.
+# Load your packages, e.g. library(drake).
+source("C.analysis/0.packages.R")  
 
+# Load our functions 
+source("C.analysis/0.func_ED.R") 
+source("C.analysis/0.func_metric.R")
 
-make(exp1_met_plan)
+# Load the drake plan 
+source("C.analysis/plan.R")      
+
+# Proces all of the outputs from ED rds files into csv files of the monthly/annual carbo fluxes for 
+# each FoRTE treatment (0, 45, 65, 85)% with the appropriate units. The files should be saved in 
+# ED-outputs/exp-constant-mon.csv and exp-constant-yr.csv 
+make(exp_constant)
+
+# Now calculate the mertic values. 
+make(metric_values)

### 2021-05-10 C.analysis/met_analysis.Rmd
Adding met analysis

---
title: "FoRTE Random Forest Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Objective

Start playing around with the met vs metric data to try in a quantiative mannner.

## Load & format data

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)

library(dplyr)
library(ggplot2)
library(tidyr)
library(MASS)
library(relaimpo)

BASE_DIR <- "/Users/dorh012/Documents/2020/FoRTE/forte-disturbance"
INPUT_DIR <- file.path(BASE_DIR, "ED-outputs", "results")

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
FORTE_SEVERITY_COLORS <- c("#000000", "#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('0 %', '45 %', '65 %', '85 %')
```

```{r}
# Load the ED output data
ED_output <- read.csv(list.files(INPUT_DIR, "exp-constant-yr.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the meteorological input, the annual average and the monthly average. 
annual_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data.csv", full.names = TRUE), stringsAsFactors = FALSE)
monthly_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data-monthly.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the results realted to the metrics of ecosystem stability, the log ratio of the time series, resilience, 
# and resistance. 
disturbance <- read.csv(file.path(INPUT_DIR, "metric", "ln_ratio_timeseries.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resilience <- read.csv(file.path(INPUT_DIR, "metric", "resilience_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resistance <- read.csv(file.path(INPUT_DIR, "metric", "resistance_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
```

```{r}
annual_met %>%  
  mutate(met = as.integer(met)) %>% 
  dplyr::select(met, met_var = variable, met_value = value) %>% 
  distinct() -> 
  met_data

resilience %>%
  left_join(resistance) %>%  
  mutate(met = as.integer(met)) %>% 
  left_join(met_data) %>%  
  dplyr::select(scn, severity, met, variable, resilience, resistance  = trough_resistance, met_var, met_value) %>%  
  dplyr::mutate(severity = as.integer(gsub(pattern = "%| ", replacement = "", x = severity))) -> 
  metrics_v_met
```

```{r}
# Prepare the data for the analysis, it should be a wide format with 
metrics_v_met %>%  
  # The year of the met data used is a categority vairable not numeric since we want ecosystem metrics of 
  # stability to compared with the met values not the met identifies (year). 
  mutate(met = as.character(met)) %>% 
  pivot_wider(names_from = met_var, values_from = met_value) -> 
  data_met_metric

#data_met_metric$severity <- factor(data_met_metric$severity, levels = c("45 %", "65 %", "85 %"), ordered = TRUE)
```

```{r}
# Seprate the resilience and resistance values into seprate data frame to make is easier to do the backwards regression. 
data_met_metric %>%  
  na.omit %>% 
  dplyr::select(-scn, -met, -resistance, -hgt) %>% 
  split(., .$variable, drop = TRUE) %>% 
  lapply(., function(x){dplyr::select(x, -variable)}) -> 
data_resilience

data_met_metric %>%  
  na.omit %>% 
  dplyr::select(-scn, -met, -resilience, -hgt) %>%  
  split(., .$variable, drop = TRUE) %>% 
  lapply(., function(x){dplyr::select(x, -variable)}) -> 
  data_resistance 
```

# Variable information 
```{r} 
annual_met %>% 
  dplyr::select(variable, description, units) %>% 
  distinct() %>%  
  knitr::kable()

```

# Step wise regressionsn

## GPP

Resistance

```{r}
# Fit the full model 
GPP_full_resistance.model <- lm(resistance ~., data = data_resistance$GPP)
summary(GPP_full_resistance.model)
```

Stepwise regression model

```{r}
# Stepwise regression model
GPP_step_resistance.model <- stepAIC(GPP_full_resistance.model, direction = "both", 
                      trace = FALSE)
summary(GPP_step_resistance.model)
```

Looking at the relative importance of the coeffiencets

```
lmg
is the R^2 contribution averaged over orderings among regressors, cf. e.g. Lindeman, Merenda and Gold 1980, p.119ff or Chevan and Sutherland (1991).

last
is each variables contribution when included last, also sometimes called usefulness.

first
is each variables contribution when included first, which is just the squared covariance between y and the variable.

pratt
is the product of the standardized coefficient and the correlation.

```

```{r}
calc.relimp(GPP_step_resistance.model, type=c("lmg","last","first","pratt"), rela=TRUE)
```

### 2021-05-10 C.analysis/paper_analysis.Rmd
Adding met analysis

@@
 ```{r}
-ggplot(data = disturbance) + 
-  geom_line(aes(year, value, color = severity, line = met), alpha = 0.5) + 
+disturbance %>%  
+  mutate(Severity = severity) %>% 
+  ggplot() + 
+  geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
   facet_wrap("variable", scales = 'free') +
   THEME +
-  labs(y = "ln(treatment/control)") +
   scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
-  labs(title = "Disturbance Response")
+  labs(y = "ln(treatment/control)", x = "Year") 
 ```

### 2021-05-29 metric_analysis.Rmd
Create metric_analysis.Rmd

---
title: "metric analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)

library(readr)
resist <- read_csv("ED-outputs/results/metric/resistance_values.csv")
met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
```

# Met data

```{r, echo=FALSE}
met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

library(ggplot2)
theme_set(theme_minimal())
ggplot(met2, aes(met, value)) +
  geom_point() + 
  facet_wrap(~variable, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90))

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)
biplot(pca_met, scale = 0)
pairs(met2_wide[-1])
```

# Metric data

```{r}
ggplot(resist, aes(trough_resistance, fill = severity)) +
  geom_histogram(bins = 20) + 
  facet_wrap(~variable)
```

# Variable importance

```{r, fig.height=8}
# Build a combined met-metric dataset
resist %>% 
  select(variable, met, trough_resistance) %>% 
  # We see from the met analysis above that vbdsf and vddsf are highly
  # correlated with nbdsf and nddsf respectively; remove one pair
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  resist_combined

resist_combined %>% 
  pivot_longer(dlwrf:vgrd) %>% 
  ggplot(aes(trough_resistance, value, color = variable)) +
  geom_point() + 
  facet_wrap(variable~name, scales="free_y") + 
  geom_smooth(method = "lm", formula = y ~ x) +
  theme(legend.position = "none", axis.text = element_text(size = 5))

# Fit a Random Forest for each variable
rc_all <- split(resist_combined, resist_combined$variable)
rf_all <- lapply(rc_all, function(x) 
  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
)
imp_all <- lapply(rf_all, function(x) {
  imp <- as.data.frame(importance(x))
  imp$met_var <- rownames(imp)
  as_tibble(imp)
})
importance <- bind_rows(imp_all, .id = "variable")
```

### 2021-05-29 metric_analysis.Rmd
Update metric_analysis.Rmd

@@
 
-# Fit a Random Forest for each variable
-rc_all <- split(resist_combined, resist_combined$variable)
+# Fit a Random Forest for each variable (across all treatment severities)
+# Note the r.f. functions don't play well with tibbles, so change to data frame
+rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
 rf_all <- lapply(rc_all, function(x) 
@@
 )
-imp_all <- lapply(rf_all, function(x) {
-  imp <- as.data.frame(importance(x))
+
+# Importance metrics and partial plot data
+import <- list()
+partial <- list()
+for(var in names(rf_all)) {
+  mod <- rf_all[[var]]
+  imp <- as.data.frame(importance(mod))
   imp$met_var <- rownames(imp)
-  as_tibble(imp)
-})
-importance <- bind_rows(imp_all, .id = "variable")
+  import[[var]] <- as_tibble(imp)
+
+  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
+  for (i in seq_along(impvar)) {
+    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
+    partial[[paste(var, i)]] <- tibble(var = var, 
+                                       met_var = impvar[i],
+                                       x = pp$x, y = pp$y)
+  }
+}
+
+import <- bind_rows(import, .id = "var")
+ggplot(import, aes(met_var, `%IncMSE`)) + geom_col() +
+  coord_flip() +
+  facet_wrap(~var)
+
+partial <- bind_rows(partial)
+ggplot(bind_rows(partial), aes(x, y, color = var)) + 
+  geom_line() + 
+  facet_wrap(~var + met_var, scales = "free")
 ```

### 2021-05-29 metric_analysis.Rmd
Explanatory text

--- 2021-05-29 metric_analysis.Rmd
@@
 # Met data
+
+Take a look at the distribution of met data, and perform a Principal Components
+Analysis (PCA) to examine auto-correlation between the various met variables.
 
@@
 
-# Variable importance
+# Met-metric data
+
+We see from the met analysis above that vbdsf and vddsf are highly
+correlated with nbdsf and nddsf respectively; remove one pair.
+
+Take a first look at the relationship between output variables and met variables.
 
@@
   geom_point() + 
-  facet_wrap(variable~name, scales="free_y") + 
+  facet_wrap(variable~name, scales = "free_y") + 
   geom_smooth(method = "lm", formula = y ~ x) +
   theme(legend.position = "none", axis.text = element_text(size = 5))
+```
 
-# Fit a Random Forest for each variable
-rc_all <- split(resist_combined, resist_combined$variable)
+# Variable importance
+
+Fit a Random Forest model to each output variable (GPP, etc.) and then calculate
+variable importance for each model. Make partial dependence plots for the
+three most important met variables for each output variable.
+
+```{r, message=FALSE}
+# Fit a Random Forest for each variable (across all treatment severities)
+# Note the r.f. functions don't play well with tibbles, so change to data frame
+library(randomForest)
+rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
 rf_all <- lapply(rc_all, function(x) 
@@
 )
-imp_all <- lapply(rf_all, function(x) {
-  imp <- as.data.frame(importance(x))
+
+# Importance metrics and partial plot data
+import <- list()
+partial <- list()
+for(var in names(rf_all)) {
+  mod <- rf_all[[var]]
+  imp <- as.data.frame(importance(mod))
   imp$met_var <- rownames(imp)
-  as_tibble(imp)
+  import[[var]] <- as_tibble(imp)
+
+  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
+  for (i in seq_along(impvar[1:3])) {
+    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
+    partial[[paste(var, i)]] <- tibble(var = var, 
+                                       met_var = impvar[i],
+                                       x = pp$x, y = pp$y)
+  }
+}
+
+import <- bind_rows(import, .id = "var")
+library(forcats)
+lapply(split(import, import$var), function(x) {
+  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
+    geom_col() +
+    coord_flip() +
+    xlab("Met variable") +
+    ggtitle(unique(x$var))
 })
-importance <- bind_rows(imp_all, .id = "variable")
+
+partial <- bind_rows(partial)
+ggplot(bind_rows(partial), aes(x, y, color = var)) + 
+  geom_line() + 
+  xlab("Met var value") + ylab("Metric") +
+  ggtitle("Partial dependence plot") +
+  facet_grid(var ~ met_var, scales = "free")
 ```
+
+# Summary
+
+Interestingly, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
+top-three most important drivers for all four output variables 
+(GPP, NEP, NPP, Rh) for resistance. Atmospheric pressure (`pres`) is
+among the top three for both GPP and NPP resistance: the former because of
+stomatal effects perhaps (?), and the latter because
+of NPP's strong dependence on GPP.

### 2021-05-30 metric_analysis.Rmd
Fix met-metric plots

@@
 library(readr)
-resist <- read_csv("ED-outputs/results/metric/resistance_values.csv")
+resistance <- read_csv("ED-outputs/results/metric/resistance_values.csv")
+resilience <- read_csv("ED-outputs/results/metric/resilience_values.csv")
 met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
@@
 met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
-pca_met <- prcomp(met2_wide[-1], scale. = TRUE)
+pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
 biplot(pca_met, scale = 0)
@@
 ```{r}
-ggplot(resist, aes(trough_resistance, fill = severity)) +
+ggplot(resistance, aes(trough_resistance, fill = severity)) +
   geom_histogram(bins = 20) + 
+  ggtitle("Resistance") +
+  facet_wrap(~variable)
+
+ggplot(resilience, aes(resilience, fill = severity)) +
+  geom_histogram(bins = 20) + 
+  ggtitle("Resilience") +
   facet_wrap(~variable)
@@
 # Build a combined met-metric dataset
-resist %>% 
+resistance %>% 
   select(variable, met, trough_resistance) %>% 
@@
   pivot_longer(dlwrf:vgrd) %>% 
-  ggplot(aes(trough_resistance, value, color = variable)) +
+  ggplot(aes(value, trough_resistance, color = variable)) +
   geom_point() + 
-  facet_wrap(variable~name, scales = "free_y") + 
-  geom_smooth(method = "lm", formula = y ~ x) +
+  facet_wrap(variable~name, scales = "free") + 
+  geom_smooth(method = "lm", formula = y ~ x, color = "black", linetype = 5) +
   theme(legend.position = "none", axis.text = element_text(size = 5))

### 2021-05-30 metric_analysis.Rmd
Analyze both resistance and resilience

@@
 
+set.seed(1234)
+
 library(dplyr)
@@
 met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
+
+resistance %>% 
+  select(scn, variable, met, value = trough_resistance, severity) %>% 
+  mutate(metric = "resistance") ->
+  combined
+resilience %>% 
+  select(scn, variable, met, value = resilience, severity) %>% 
+  mutate(metric = "resilience") %>% 
+  bind_rows(combined) ->
+  combined
 ```
@@
 
-# Met-metric data
+# Met-metric data {.tabset}
 
@@
 
+## Resistance
+
 ```{r, fig.height=8}
 # Build a combined met-metric dataset
-resistance %>% 
-  select(variable, met, trough_resistance) %>% 
+combined %>% 
+  select(metric, variable, met, value) %>% 
   # We see from the met analysis above that vbdsf and vddsf are highly
@@
   left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
+  metric_combined
+
+resistance %>% 
+  select(variable, met, trough_resistance) %>% 
+  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
   resist_combined
 
-resist_combined %>% 
-  pivot_longer(dlwrf:vgrd) %>% 
-  ggplot(aes(value, trough_resistance, color = variable)) +
+metric_combined %>% 
+  pivot_longer(dlwrf:vgrd, values_to = "met_value") ->
+  metric_comb_long
+
+p_mm <- ggplot(filter(metric_comb_long, metric == "resistance"), 
+            aes(met_value, value, color = variable)) +
   geom_point() + 
-  facet_wrap(variable~name, scales = "free") + 
   geom_smooth(method = "lm", formula = y ~ x, color = "black", linetype = 5) +
+  facet_wrap(~ variable + name, scales = "free") + 
+  ggtitle("Resistance") +
   theme(legend.position = "none", axis.text = element_text(size = 5))
+print(p_mm)
 ```
 
-# Variable importance
+## Resilience
+
+```{r, fig.height=8}
+p_mm <- p_mm %+% filter(metric_comb_long, metric == "resilience")
+print(p_mm + ggtitle("Resilience"))
+```
+
+# Variable importance {.tabset}
 
@@
 library(randomForest)
-rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
+rc_all <- split(as.data.frame(metric_combined), ~ metric + variable)
 rf_all <- lapply(rc_all, function(x) 
-  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
+  # remove metric, variable, met columns and run RF
+  randomForest(value ~ ., data = x[-1:-3], importance = TRUE)
 )
@@
   mod <- rf_all[[var]]
+  metric <- rc_all[[var]]$metric[1]
+  variable <- rc_all[[var]]$variable[1]
+
   imp <- as.data.frame(importance(mod))
   imp$met_var <- rownames(imp)
+  imp$metric <- metric
+  imp$variable <- variable
   import[[var]] <- as_tibble(imp)
@@
     pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
-    partial[[paste(var, i)]] <- tibble(var = var, 
+    partial[[paste(var, i)]] <- tibble(metric = metric,
+                                       variable = variable,
                                        met_var = impvar[i],
@@
 
-import <- bind_rows(import, .id = "var")
+import <- bind_rows(import)
+partial <- bind_rows(partial)
+```
+
+## Resistance
+
+```{r}
 library(forcats)
-lapply(split(import, import$var), function(x) {
+f_imp_plot <- function(x) {
   ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
@@
     xlab("Met variable") +
-    ggtitle(unique(x$var))
-})
+    ggtitle(unique(x$variable))
+}
+lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)
 
-partial <- bind_rows(partial)
-ggplot(bind_rows(partial), aes(x, y, color = var)) + 
+p_pdp <- ggplot(filter(partial, metric == "resistance"), 
+                aes(x, y, color = variable)) + 
   geom_line() + 
-  xlab("Met var value") + ylab("Metric") +
+  xlab("Met var value") + ylab("Resistance") +
   ggtitle("Partial dependence plot") +
-  facet_grid(var ~ met_var, scales = "free")
+  facet_grid(variable ~ met_var, scales = "free")
+print(p_pdp)
+```
+
+## Resilience
+
+```{r}
+# Re-do above plots but now for resilience
+lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
+p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
+print(p_pdp + ylab("Resilience"))
 ```
@@
 
-Interestingly, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
+For **resistance**, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
 top-three most important drivers for all four output variables 
-(GPP, NEP, NPP, Rh) for resistance. Atmospheric pressure (`pres`) is
-among the top three for both GPP and NPP resistance: the former because of
-stomatal effects perhaps (?), and the latter because
-of NPP's strong dependence on GPP.
+(GPP, NEP, NPP, Rh). Higher temperature and more light translate into higher
+resistance for all variables.
+
+This is largely true for **resilience** as well, although
+`prate` (precipitation) is important for both NPP and Rh; in fact precipitation
+is far and away the most important variable for Rh. Here, however,
+higher temperature and more light mean _less_ resilience.

### 2021-06-08 C.analysis/paper_analysis.Rmd
Update manuscript figs

--- 2021-05-06 C.analysis/paper_analysis.Rmd
@@
 library(ggplot2)
+library(stringr)
 
@@
 
-```{r}
+```{r, fig.width=6.5, fig.height=4, fig.retina = 4}
 ED_output %>%  
   filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
-  mutate(description = paste0(variable, ' (', unit,  ')')) %>% 
-  ggplot(aes(year, value, color = met, group = met)) + 
+  mutate(description = paste0(variable, '\n (', unit,  ')')) %>% 
+  ggplot(aes(year, value, group = met,  color = met)) + 
   geom_line() + 
-   facet_wrap("description", scales = 'free', strip.position = "top",
-             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
+  facet_wrap("description", scales = "free",  
+             labeller = label_glue('({.l}) {description}')) + 
   THEME +
-  labs(y = "")
+  labs(y = "", x = "Year")  + 
+  theme(legend.position = "none") -> 
+  plot
+
+plot + 
+  ggsave("figures/baseline.png", width = 6, height = 4.5)
+
 ```
@@
 ```{r}
-ggplot(data = disturbance) + 
-  geom_line(aes(year, value, color = severity, line = met), alpha = 0.5) + 
-  facet_wrap("variable", scales = 'free') +
+disturbance %>%  
+  mutate(Severity = severity) %>% 
+  ggplot() + 
+  geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
+  facet_wrap("variable", scales = 'free', 
+             labeller = label_glue('({.l}) {variable}')) +
   THEME +
-  labs(y = "ln(treatment/control)") +
   scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
-  labs(title = "Disturbance Response")
+  guides(colour = guide_legend("Severity")) + 
+  labs(y = "Log Ratio", x = "Years Since Disturbance") -> 
+  plot
+
+plot
+plot + 
+  ggsave("figures/disturbance.png", width = 6, height = 4.5)
+
 ```
@@
 
-```{r}
+```{r, fig.width=5, fig.height=5}
 ggplot(data = resistance) + 
-  geom_point(aes(severity, trough_resistance, color = severity), position = position_jitter(), alpha = 0.5, size = 2) + 
-  facet_wrap("variable", scales = "free") + 
-  labs(y = "ln(dist/control) unitless", 
-       title = "Resistance") + 
-  THEME + 
-  scale_color_manual(values = FORTE_SEVERITY_COLORS)
+  geom_point(aes(severity, trough_resistance, color = severity),
+             position = position_jitter(), 
+             alpha = 0.75, size = 1.5) + 
+  facet_wrap("variable", scales = "free",  
+             labeller = label_glue('({.l}) {variable}')) + 
+  labs(y = "Log Ratio", 
+       title = "Resistance", x = "Severity") + 
+  theme_bw(base_size = 10) +
+  guides(colour = guide_legend("Severity")) + 
+  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
+  plot
+
+plot + 
+  ggsave("figures/resistance.png", width = 6, height = 4.5)
 
@@
 
-```{r}
+```{r, fig.width=5, fig.height=5}
 ggplot(data = resilience) + 
-  geom_point(aes(severity, resilience, color = severity), position = position_jitter(), alpha = 0.5, size = 2) + 
-  facet_wrap("variable", scales = "free") + 
-  labs(y = "ln(dist/control) unitless", 
-       title = "Resilience") + 
-  THEME + 
-  scale_color_manual(values = FORTE_SEVERITY_COLORS)
+  geom_point(aes(severity, resilience, color = severity),              position = position_jitter(), 
+             alpha = 0.75, size = 1.5) + 
+  facet_wrap("variable", scales = "free",  
+             labeller = label_glue('({.l}) {variable}')) + 
+  labs(y = "Log Ratio", 
+       title = "Resilience", 
+        x = "Severity") + 
+  theme_bw(base_size = 10) +
+  guides(colour = guide_legend("Severity")) + 
+  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
+  plot
+
+plot + 
+  ggsave("figures/resilience.png", width = 6, height = 4.5)
 ```
@@
 
-```{r}
+```{r, fig.width=5, fig.height=5}
 resilience %>% 
@@
   ggplot(aes(trough_resistance, resilience, color = severity)) + 
-  geom_point(alpha = 0.5, size = 2) + 
-   facet_wrap("variable", scales = "free") + 
-  labs(x = "resistance", 
-       y = "resilience", 
+  geom_point(alpha = 0.75, size = 1.5) + 
+  facet_wrap("variable", scales = "free",  
+             labeller = label_glue('({.l}) {variable}')) + 
+  labs(x = "Resistance", 
+       y = "Resilience", 
        title = "Resilience vs. Resistance") + 
-  THEME + 
-  scale_color_manual(values = FORTE_SEVERITY_COLORS)
+  theme_bw(base_size = 10) +
+  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
+  guides(colour = guide_legend("Severity")) -> 
+  plot
+
+plot + 
+  ggsave("figures/resistance_resilience.png", width = 6, height = 4.5)
+
 ```
@@
 
-```{r, fig.width=8, fig.height=8}
+```{r, fig.width=5, fig.height=5}
 annual_met %>% 
@@
 
-```{r, fig.width=8, fig.height=8}
+```{r, fig.width=10, fig.height=5.5}
 monthly_met %>% 
-  mutate(description = paste0(description, ' (', units, ')')) %>% 
+  mutate(units = if_else(variable == "vgrd", 'm s-1', units)) %>% 
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
   ggplot(aes(mon_time, value, color = met, group = met, line = met)) + 
@@
    facet_wrap("description", scales = 'free', strip.position = "top",
-             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
+             labeller = label_wrap_gen(width = 20)) +
+  labs(y = NULL, x  = NULL) +
   THEME +
-  labs(y = NULL, x  = NULL, title = "Monthly Met") + 
-  theme(legend.position = "none")
+  scale_x_continuous(labels = c("Jan", "Jun", "Dec"), breaks = c(1, 6, 12)) +
+  theme(legend.position = "none") -> plot
+
+plot + 
+  ggsave("figures/monthly_met.png", width = 10, height = 10)
+
 ```

### 2021-06-08 C.analysis/0.func_metric.R

@@ -109,7 +109,7 @@ fit_data <- function(d){
 # Extract the slope from the list of teh fits 
 # Args 
 #   l: a list of the data frames 
-#   n: the list of the slopes freates by fit_data 
+#   n: the list of the slopes names created by fit_data 
 # Return: a data frame of the slope fits
 extract_fit_data <- function(l, n){
   
@@ -117,15 +117,18 @@ extract_fit_data <- function(l, n){
   
   # If it was determined that there was no segment
   if(length(l[["seg.mod"]]) == 12){
-    seg_slope1 <- l[["seg.mod"]]$coefficients[2]
+    seg_slope1 <- ln_slope
+    temporal_stability <- sigma(l[["lin.mod"]])
     seg_slope2 <- NA
     seg_brk <- NA 
     
   } else {
     slopes <- slope(l[["seg.mod"]])
     seg_slope1 <- slopes$x[1,1] # Extract the slope of the first segment
+    temporal_stability <- sigma(l[["seg.mod"]]) # Etract the standard error about the
     seg_slope2 <- slopes$x[2,1] # Extract the slope of the second segment
     seg_brk <- l[["seg.mod"]]$psi[1,3] # Extract the year of the break point 
+    
   }
   
   # Extract the data about the fit. 
@@ -138,8 +141,10 @@ extract_fit_data <- function(l, n){
                    lin_slope = ln_slope, 
                    seg_slope1 = seg_slope1, 
                    seg_slope2 = seg_slope2, 
-                   seg_brk = seg_brk) 
+                   seg_brk = seg_brk, 
+                   temporal_stability = temporal_stability) 
   
+  return(dt)
 }
 
 
@@ -179,6 +184,39 @@ get_resilience <- function(d){
     select(scn, severity, met, variable, resilience = seg_slope1) %>% 
     # TODO figure out where this is coming from 
     filter(!is.na(resilience))
+}
+
+get_temporal_stability <- function(d){
+  
+  d <- as.data.table(d)
+  disturbance <- calculate_ln_ratio(d=d)
+  resistance <- get_through_resistance(d=d)
+  
+  # Get the data that will be used to calculate the resilience.
+  # We will only use the data from after the 
+  disturbance %>% 
+    left_join(select(resistance, met, variable, severity, trough_year),     
+              by = c("variable", "met", "severity")) %>% 
+    dplyr::filter(year >= trough_year) %>% 
[…]

### 2021-06-08 C.analysis/plan.R

@@ -64,7 +64,8 @@ metric_values <- drake_plan(
     data <- get_resilience(ED_data)
     fname <- file.path(metric_dir, 'resilience_values.csv')
     write.csv(data, file = fname, row.names = FALSE)
-  })
+  }), 
+  
 )
 
 

### 2021-06-28 C.analysis/paper_analysis.Rmd
pre co-author feedback

@@
 library(stringr)
+library(stickylabeller)
+library(ggpmisc)
+library(ggpubr)
 
@@
 # Baseline Scenario
+
+```{r}
+ED_output %>%  
+  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP", "AGB")) %>% 
+  mutate(description = paste0(variable, '\n (', unit,  ')')) %>% 
+  dplyr::filter(year == 2020) %>% 
+  group_by(variable, unit) %>% 
+  summarise(sd = sd(value), mean = mean(value)) %>% 
+  ungroup %>% 
+  mutate(cv = sd/mean)
+
+```
 
@@
 
+```{r}
+
+resistance %>% 
+  group_by(variable, severity) %>%  
+  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance))
+
+```
+
 ```{r, fig.width=5, fig.height=5}
@@
 # Resilience
+
+```{r}
+resilience %>% 
+  group_by(variable, severity) %>% 
+  summarise(mean = mean(resilience))
+
+```
 
@@
   geom_point(alpha = 0.75, size = 1.5) + 
-  facet_wrap("variable", scales = "free",  
+  facet_wrap("variable", #scales = "free",  
              labeller = label_glue('({.l}) {variable}')) + 
@@
   ggsave("figures/resistance_resilience.png", width = 6, height = 4.5)
+
+```
+
+```{r, fig.width=5, fig.height=5}
+my.formula <- y ~ x
+
+resilience %>% 
+  left_join(resistance) %>%  
+  ggplot(aes(trough_resistance, resilience, color = severity)) + 
+  geom_point(alpha = 0.75, size = 1.5) + 
+  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
+  stat_cor(method = "pearson", aes(label = ..r.label..),
+           size = 3, label.x.npc = "left", label.y.npc = "bottom", 
+           r.accuracy = 0.01) + 
+  labs(x = "Resistance", 
+       y = "Resilience", 
+       title = "Resilience vs. Resistance") + 
+  theme_bw(base_size = 10) +
+  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
+    facet_wrap("variable", scales = "free",  
+            labeller = label_glue('({.l}) {variable}')) + 
+  guides(colour = guide_legend("Severity")) +
+  ggsave("figures/resistance_resilience3.png", width = 6, height = 4.5)
+
+```
+
+```{r}
+
+resilience %>% 
+  left_join(resistance) %>%  
+  ggplot(aes(trough_resistance, resilience, color = severity)) + 
+  geom_point(alpha = 0.75, size = 1.5) + 
+  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
+  stat_cor(method = "pearson", aes(label = ..r.label..),
+           size = 3, label.x.npc = "left", label.y.npc = "bottom", 
+           r.accuracy = 0.01) + 
+  labs(x = "Resistance", 
+       y = "Resilience", 
+       title = "Resilience vs. Resistance") + 
+  theme_bw(base_size = 10) +
+  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
+  guides(colour = guide_legend("Severity"))
 

### 2021-07-09 metric_analysis.Rmd

@@
 }
-lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)
+xx <- lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)
+
+library(cowplot)
+varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
+print(varimp_plot)
+save_plot("varimp_resistance.png", varimp_plot)
 
@@
 # Re-do above plots but now for resilience
-lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
+xx <- lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
+varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
+print(varimp_plot)
+save_plot("varimp_resilience.png", varimp_plot)
+
 p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
@@
 # Summary
+
+```{r}
+# Combined partial dependence plot
+partial %>% 
+  mutate(metric = if_else(metric == "resistance", "Resistance", "Resilience")) %>% 
+  ggplot(aes(x, y, color = variable)) + 
+  geom_line() + 
+  xlab("Meteorological variable value") + ylab("Metric value") +
+  scale_color_discrete("") +
+  facet_grid(metric ~ met_var, scales = "free") +
+  theme(axis.text.x = element_text(angle = 90))
+```
 

### 2021-07-27 C.analysis/paper_analysis.Rmd
Update to manuscript circualted to co-authors at end of July

@@
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)
-
+# R 3.6.3
 library(dplyr)
@@
 THEME <- theme_bw()
-FORTE_SEVERITY_COLORS <- c("#000000", "#009E73", "#0072B2", "#D55E00")
-names(FORTE_SEVERITY_COLORS) <- c('0 %', '45 %', '65 %', '85 %')
+FORTE_SEVERITY_COLORS <- c("#009E73", "#0072B2", "#D55E00")
+names(FORTE_SEVERITY_COLORS) <- c('45 %', '65 %', '85 %')
 ```
@@
 
+```{r}
+annual_met %>% 
+  filter(variable %in% c("ugrd", "vgrd")) %>% 
+  mutate(value = value^2) %>% 
+  group_by(year, met) %>% 
+  summarise(value = sqrt(sum(value))) %>% 
+  ungroup() %>%
+  mutate(units = "m s-1", 
+         description = "total winnd", 
+         variable = "grd") -> 
+  wind_df
+
+annual_met %>% 
+  filter(!variable %in% c("ugrd", "vgrd")) %>% 
+  bind_rows(wind_df) %>% 
+  filter(variable != "hgt") ->
+  annual_met
+
+monthly_met %>% 
+  filter(variable %in% c("ugrd", "vgrd")) %>% 
+  mutate(value = value^2) %>% 
+  group_by(month, mon_time, met) %>% 
+  summarise(value = sqrt(sum(value))) %>% 
+  ungroup() %>%
+  mutate(units = "m s-1", 
+         description = "total winnd", 
+         variable = "grd") -> 
+  wind_df
+
+monthly_met %>% 
+  filter(!variable %in% c("ugrd", "vgrd")) %>% 
+  bind_rows(wind_df) %>% 
+  filter(variable != "hgt") ->
+  monthly_met
+```
+
 # Baseline Scenario
@@
 ```{r, fig.width=6.5, fig.height=4, fig.retina = 4}
+
+# # UMBS forest observations taken from table 2 of 
+# C.M. Gough, C.S. Vogel, H.P. Schmid, H.-B. Su, P.S. Curtis,
+# Multi-year convergence of biometric and meteorological estimates of forest carbon storage,
+# Agricultural and Forest Meteorology,
+# Volume 148, Issue 2,
+# 2008,
+# Pages 158-170,
+# ISSN 0168-1923,
+# https://doi.org/10.1016/j.agrformet.2007.08.004
+data.frame(description = c("NPP\n (MgC ha-1 year-1)", "NEP\n (MgC ha-1 year-1)"), 
+           mean = c(6.54, 1.53), 
+           lower =  c(6.54 - .76, 1.53 - 1.15), 
+           upper = c(6.54 + .76,1.53 + 1.15 ), 
+           year = c(2001, 2001)) -> 
+  obs
+
 ED_output %>%  
   filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
-  mutate(description = paste0(variable, '\n (', unit,  ')')) %>% 
-  ggplot(aes(year, value, group = met,  color = met)) + 
-  geom_line() + 
+  mutate(description = paste0(variable, '\n (', unit,  ')')) -> 
+  to_plot
+
+  ggplot() + 
+  geom_line(data = to_plot, aes(year, value, group = met), 
+            alpha = 0.9, color = "#999999", ) + 
+  geom_pointrange(data = obs, aes(x=year, y=mean, ymin=lower, ymax=upper)) + 
   facet_wrap("description", scales = "free", 
@@
 
-plot + 
-  ggsave("figures/baseline.png", width = 6, height = 4.5)
+plot
+ggsave(plot, filename = "figures/baseline.png", width = 6, height = 4.5)
 
@@
   geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
-  facet_wrap("variable", scales = 'free', 
-             labeller = label_glue('({.l}) {variable}')) +
+  facet_wrap("variable", scales = 'free') +
+            # labeller = label_glue('({.l}) {variable}')) +
   THEME +
@@
 plot
-plot + 
-  ggsave("figures/disturbance.png", width = 6, height = 4.5)
+
+ggsave(plot, filename = "figures/disturbance.png", width = 6, height = 4.5)
 
@@
   group_by(variable, severity) %>%  
-  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance))
+  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance)) %>% 
+  ungroup(.) -> 
+  summary_table
+```
+
+```{r}
+summary_table %>% 
+  select(variable, severity, mean) %>% 
+  tidyr::spread(severity, mean) %>% 
+  mutate(change1 = `65 %` - `45 %` / `45 %`, 
+         change2 = `85 %` - `65 %` / `65 %`)
 
@@
 ```{r, fig.width=5, fig.height=5}
-ggplot(data = resistance) + 
-  geom_point(aes(severity, trough_resistance, color = severity),
+ggplot() + 
+  geom_point(data = resistance, aes(severity, trough_resistance, color = severity),
              position = position_jitter(), 
              alpha = 0.75, size = 1.5) + 
-  facet_wrap("variable", scales = "free",  
+    geom_pointrange(data = summary_table, 
+                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
+                  color = "black") + 
+    ylim(-1.5, 0) +
+  facet_wrap("variable",
              labeller = label_glue('({.l}) {variable}')) + 
-  labs(y = "Log Ratio", 
-       title = "Resistance", x = "Severity") + 
+  labs(y = "Resistance", 
+       x = "Severity") + 
   theme_bw(base_size = 10) +
-  guides(colour = guide_legend("Severity")) + 
+  theme(legend.position = "none") +
   scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
@@
 
-plot + 
-  ggsave("figures/resistance.png", width = 6, height = 4.5)
-
+ggsave(plot, filename = "figures/resistance.png", width = 6, height = 4.5)
 ```
@@
   group_by(variable, severity) %>% 
-  summarise(mean = mean(resilience))
+  summarise(mean = mean(resilience), sd = sd(resilience)) %>% 
+  ungroup(.) -> 
+  summary_table
+```
+
+```{r}
+summary_table %>% 
+  select(variable, severity, mean) %>% 
+  tidyr::spread(severity, mean) %>% 
+  mutate(change1 = `65 %` - `45 %` / `45 %`, 
+         change2 = `85 %` - `65 %` / `65 %`)
 
@@
 ```{r, fig.width=5, fig.height=5}
-ggplot(data = resilience) + 
-  geom_point(aes(severity, resilience, color = severity),              position = position_jitter(), 
+ggplot() + 
+  geom_point(data = resilience, aes(severity, resilience, color = severity),              
+             position = position_jitter(), 
              alpha = 0.75, size = 1.5) + 
-  facet_wrap("variable", scales = "free",  
-             labeller = label_glue('({.l}) {variable}')) + 
-  labs(y = "Log Ratio", 
-       title = "Resilience", 
+  facet_wrap("variable", labeller = label_glue('({.l}) {variable}')) + 
+  geom_pointrange(data = summary_table, 
+                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
+                    color = "black") + 
+  labs(y = "Resilience", 
         x = "Severity") + 
+  ylim(0, 0.035)+
   theme_bw(base_size = 10) +
-  guides(colour = guide_legend("Severity")) + 
+  theme(legend.position = "none") +
   scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
@@
 
-plot + 
-  ggsave("figures/resilience.png", width = 6, height = 4.5)
+  ggsave(plot, filename = "figures/resilience.png", width = 6, height = 4.5)
 ```
@@
   geom_point(alpha = 0.75, size = 1.5) + 
-  facet_wrap("variable", #scales = "free",  
-             labeller = label_glue('({.l}) {variable}')) + 
+  facet_wrap("variable")+ #scales = "free",  
+            # labeller = label_glue('({.l}) {variable}')) + 
   labs(x = "Resistance", 
@@
 
-plot + 
-  ggsave("figures/resistance_resilience.png", width = 6, height = 4.5)
+ggsave(plot, filename = "figures/resistance_resilience.png", width = 6, height = 4.5)
 
@@
   geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
-  stat_cor(method = "pearson", aes(label = ..r.label..),
-           size = 3, label.x.npc = "left", label.y.npc = "bottom", 
+  stat_cor(method = "pearson", 
+           aes(label = ..r.label..),
+           size = 3, 
+           label.y.npc = "bottom",
            r.accuracy = 0.01) + 
@@
   scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
-    facet_wrap("variable", scales = "free",  
+    facet_wrap("variable", 
+               scales = "free", 
                labeller = label_glue('({.l}) {variable}')) + 
-  guides(colour = guide_legend("Severity")) +
-  ggsave("figures/resistance_resilience3.png", width = 6, height = 4.5)
+  guides(colour = guide_legend("Severity")) -> 
+  plot
+  ggsave(plot, filename = "figures/resistance_resilience3.png", width = 6, height = 4.5)
 
@@
   mutate(units = if_else(variable == "vgrd", 'm s-1', units)) %>% 
-  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
-  ggplot(aes(mon_time, value, color = met, group = met, line = met)) + 
-  geom_line() +
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
+  data
+
+data %>% 
+  filter(met == "2019") -> 
+  data_2019
+
+ggplot() + 
+  geom_line(data = data, aes(mon_time, value, group = met, line = met), 
+            alpha = 0.8, color = "#999999") +
+  geom_line(data = data_2019, aes(mon_time, value, color = met, group = met, line = met), 
+            color = "#D55E00", size = 1) +
    facet_wrap("description", scales = 'free', strip.position = "top",
-             labeller = label_wrap_gen(width = 20)) +
+             labeller = label_wrap_gen(width = 20), ncol = 2) +
   labs(y = NULL, x  = NULL) +
@@
   scale_x_continuous(labels = c("Jan", "Jun", "Dec"), breaks = c(1, 6, 12)) +
-  theme(legend.position = "none") -> plot
+  theme(legend.position = "none") -> 
+  plot
 
-plot + 
-  ggsave("figures/monthly_met.png", width = 10, height = 10)
+plot
 
+ggsave(plot, filename = "figures/monthly_met.png", width = 6, height = 10)
 ```

### 2021-07-27 C.analysis/get_metrics.R

@@ -1,25 +1,0 @@
-# Load helper functions. 
-source(here::here('C.analysis', 'packages.R'))
-source(here::here('C.analysis', 'ED_functions.R'))
-source(here::here('C.analysis', 'metric_functions.R'))
-
-# Define the inputs and output files. 
-ED_OUTPUT_DIR <- here::here('ED-outputs')
-WRITE_TO      <-  here::here('ED-outputs', 'results')
-dir.create(WRITE_TO, showWarnings = FALSE)
-
-# The carbon fluxes to process. 
-carbon_flux_var_names <- c("GPP", "NPP", "Rh", "NEP")
-
-# Load the formated ED data. 
-file.path(ED_OUTPUT_DIR, "results", "exp-constant-yr.csv") %>%  
-  read.csv(stringsAsFactors = FALSE) %>% 
-  dplyr::filter(variable %in% carbon_flux_var_names) -> 
-  ED_data
-
-# Get the time series of the log ratio aka the disturbance. 
-ln_ratio_ts <- calculate_ln_ratio(ED_data)
-
-# Calculate our metrics of stability.
-resistance <- get_through_resistance(ED_data)
-resilience <- get_resilience(ED_data)

### 2021-08-18 metric_analysis.Rmd

--- 2021-05-30 metric_analysis.Rmd
@@
 library(dplyr)
+library(stringr)
+library(readr)
+resistance <- read_csv("./ED-outputs/results/metric/resistance_values.csv")
+resilience <- read_csv("./ED-outputs/results/metric/resilience_values.csv")
+met <- read_csv("./ED-outputs/results/constant_annual_met_data.csv")
 
-library(readr)
-resistance <- read_csv("ED-outputs/results/metric/resistance_values.csv")
-resilience <- read_csv("ED-outputs/results/metric/resilience_values.csv")
-met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
+# The variable table, contains inforamtion about the met variable abreviation with a descrioption. 
+var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
+                                    'vddsf', 'prate', 'pres', 'hgt', 
+                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP', 'grd'), 
+                       description = c('downward long wave radiation', 
+                                       'near infrared beam downward solar radiation', 
+                                       'near IR diffuse downward solar radiation', 
+                                       'visible beam downward solar radiation',
+                                       'visible diffuse downward solar radiation', 
+                                       'precipitation rate', 
+                                       'atmospheric pressure', 
+                                       'geopotential height', 
+                                       'zonal wind',
+                                       'meridional wind', 
+                                       'specific humidity',
+                                       'air temperature', 
+                                       'pre-dist. productivity', 'total wind'),  
+                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
+                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
+                                 'm s-1', 'm s-1', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1', 'm s-1'), stringsAsFactors=FALSE)
+
+# Add 2000 GPP to the met df since it is a "condition". 
+ED_output <- read.csv("ED-outputs/results/exp-constant-yr.csv", stringsAsFactors = FALSE) %>% 
+  filter(scn == "harvest_0" & variable == "GPP" & year == 2000) %>% 
+  select(met, value, variable, value, units = unit)
+
+met %>% 
+  filter(variable %in% c("ugrd", "vgrd")) %>% 
+  mutate(value = value^2) %>% 
+  group_by(year, met) %>% 
+  summarise(value = sqrt(sum(value))) %>% 
+  ungroup() %>%
+  mutate(units = "m s-1", 
+         description = "total winnd", 
+         variable = "grd") -> 
+  wind_df
+
+met %>% 
+  filter(!variable %in% c("ugrd", "vgrd")) %>% 
+  bind_rows(wind_df, 
+            ED_output) ->
+  met
 
@@
 ```{r, echo=FALSE}
+des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
+               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
+               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
+                "Precipitation Rate\n(kgH2O m-2 s-1)", "Total Wind\n(m s-1)", 
+               "Pre-Dist. Productivity\n(MgC ha-1 year-1)")
+
 met %>% 
@@
 
+met2 %>% 
+  mutate(met_var = variable) %>% 
+  inner_join(var_info, by = "met_var") %>%  
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
+  filter(description %in% des_order) -> 
+  to_plot_met2
+
+met2 %>% 
+  filter(met == 2019) -> 
+  met2019
+
+met2019 %>% 
+  mutate(met_var = variable) %>% 
+  left_join(var_info, by = "met_var") %>%  
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
+  to_plot_2019
+
+to_plot_met2$description <- factor(to_plot_met2$description, des_order, 
+                              ordered = TRUE) 
+to_plot_2019$description <- factor(to_plot_2019$description, des_order, 
+                              ordered = TRUE)
+
 library(ggplot2)
 theme_set(theme_minimal())
-ggplot(met2, aes(met, value)) +
+ggplot(to_plot_met2, aes(met, value)) +
   geom_point() + 
-  facet_wrap(~variable, scales = "free_y") +
-  theme(axis.text.x = element_text(angle = 90))
+  #geom_point(data = met2019, aes(met, value), color = "#D55E00") +
+  facet_wrap(~description, scales = "free_y", labeller = label_wrap_gen(width = 20)) +
+  theme(axis.text.x = element_text(angle = 90)) + 
+  labs(x = NULL, y = 'Mean annual conditions value') -> 
+  plot
+
+ggsave(plot, filename = "./C.analysis/figures/conditions_yr.png", width = 10, height = 6)
+plot
 
@@
 pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
-biplot(pca_met, scale = 0)
+bit(pca_met, scale = 0)
 pairs(met2_wide[-1])
@@
 metric_combined %>% 
-  pivot_longer(dlwrf:vgrd, values_to = "met_value") ->
+  pivot_longer(dlwrf:vddsf, values_to = "met_value") ->
   metric_comb_long
@@
 library(randomForest)
-rc_all <- split(as.data.frame(metric_combined), ~ metric + variable)
+d <- as.data.frame(metric_combined)
+rc_all <- split(d, interaction(d$metric, d$variable))
 rf_all <- lapply(rc_all, function(x) 
@@
 ```{r}
+
 library(forcats)
@@
     geom_col() +
+    scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
     coord_flip() +
-    xlab("Met variable") +
+    xlab(NULL) +
     ggtitle(unique(x$variable)) 
 }
-lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)
+
+import %>%
+  left_join(var_info) %>% 
+  mutate(met_var = str_to_title(description)) -> 
+  import2
+
+d <- filter(import2, metric == "resistance")
+xx <- lapply(split(d, d$variable), f_imp_plot)
+
+library(cowplot)
+varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
+print(varimp_plot)
+
+ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resistance.png", width = 10, height = 6)
+
+save_plot("varimp_resistance.png", varimp_plot)
 
@@
 # Re-do above plots but now for resilience
-lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
+d <- filter(import2, metric == "resilience")
+xx <- lapply(split(d, d$variable), f_imp_plot)
+varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
+print(varimp_plot)
+ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resilience.png", width = 10, height = 6)
+
+save_plot("varimp_resilience.png", varimp_plot)
+
 p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
@@
 # Summary
+
+```{r}
+
+var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
+                                    'vddsf', 'prate', 'pres', 'hgt', 
+                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP'), 
+                       description = c('downward long wave radiation', 
+                                       'near infrared beam downward solar radiation', 
+                                       'near IR diffuse downward solar radiation', 
+                                       'visible beam downward solar radiation',
+                                       'visible diffuse downward solar radiation', 
+                                       'precipitation rate', 
+                                       'atmospheric pressure', 
+                                       'geopotential height', 
+                                       'zonal wind',
+                                       'meridional wind', 
+                                       'specific humidity',
+                                       'air temperature', 
+                                       'pre-dist. productivity'),  
+                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
+                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
+                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1'), stringsAsFactors=FALSE)
+
+var_info$met_var <- factor(var_info$met_var, unique(var_info$met_var), ordered = TRUE)
+
+partial %>% 
+  left_join(var_info) %>% 
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
+  data_partial
+
+# Combined partial dependence plot
+data_partial %>% 
+  mutate(metric = if_else(metric == "resistance", "Resistance", "Resilience")) ->  
+  to_plot
+
+des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
+               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
+               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
+                "Precipitation Rate\n(kgH2O m-2 s-1)", "Pre-Dist. Productivity\n(MgC ha-1 year-1)")
+
+to_plot$description <- factor(to_plot$description, des_order, 
+                              ordered = TRUE)
+
+to_plot %>% 
+  ggplot(aes(x, y, color = variable)) + 
+  geom_line() + 
+  xlab("Conditions variable value") + ylab("Metric value") +
+  scale_color_discrete("") +
+  #facet_grid(metric ~ description, scales = "free", switch = "y") +
+  facet_grid(metric ~ description, 
+             labeller = label_wrap_gen(width = 10), scales = "free", switch = "y") +
+  theme_bw()+ 
+  theme(axis.text.x = element_text(angle = 90)) -> 
+  plot
+
+plot
+
+ggsave(plot, filename = "./C.analysis/figures/cond_metric.png", width = 10, height = 6)
+```
 
@@
 higher temperature and more light mean _less_ resilience.
+
+Interesting that when GPP is included as a condition that could explain the resistance & resilience metrics it really only affect the Rh values.
+
+```{r}
+met %>% 
+  filter(variable == "GPP") %>% 
+  select(met, pre_GPP = value) -> 
+  pre_dist_prod
+
+met %>% 
+  filter(!variable %in% c("GPP", "hgt")) %>% 
+  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
+  select(met, value, description) -> 
+  met_values
+
+equal_breaks <- function(n = 3, s = 0.05, ...){
+  function(x){
+    # rescaling
+    d <- s * diff(range(x)) / (1+2*s)
+    signif(seq(min(x)+d, max(x)-d, length=n), digits = 3)
+  }
+}
+
+met_values %>% 
+  left_join(pre_dist_prod) %>% 
+  ggplot(aes(value, pre_GPP)) + 
+  geom_point() +
+  facet_wrap(~description, scales = "free_x", labeller = label_wrap_gen(width = 20)) + 
+  labs(y = "Pre-disturbance productivity (MgC ha-1)", 
+       x = "Meteorological value") + 
+  scale_x_continuous(breaks=equal_breaks(n=3, s=0.05), 
+                     expand = c(0.05, 0)) + 
+    theme(axis.text.x = element_text(angle=90)) -> 
+  plot
+
+ggsave(plot, filename = "./C.analysis/figures/gpp_met.png", width = 10, height = 6)
+
+```

### 2021-10-29 metric_analysis.Rmd
Minor revisions from JGR

@@
   mutate(units = "m s-1", 
-         description = "total winnd", 
+         description = "total wind", 
          variable = "grd") -> 

### 2021-10-29 C.analysis/paper_analysis.Rmd

@@
 
+chrono <- rbind(
+  data.frame(value = c(0.55, 1.35, 1.30, 0.80, 0.45, 1.53), 
+             year = c(6, 24, 50, 56, 68, 85), 
+           variable = "NEP", 
+           description = "NEP\n (MgC ha-1 year-1)"), 
+  data.frame(value = c(5.07, 5.13, 5.44, 5.27,  5.04, 6.54),
+             year = c(6, 24, 50, 56, 68, 85), 
+             variable = "NPP",        
+             description = "NPP\n (MgC ha-1 year-1)"), 
+    data.frame(value = c(4.52, 3.78, 4.14, 4.47, 4.59, 5.01),
+             year = c(6, 24, 50, 56, 68, 85), 
+             variable = "Rh",        
+             description = "Rh\n (MgC ha-1 year-1)"))
+
+chrono$year <- chrono$year + 1900
+
+# read in other UMBS data 
+read.csv(file.path(BASE_DIR, "C.analysis", "UMBS_data",
+                   "US-UMB_AnnualNEP_2002-2019.csv")) %>%
+  tidyr::gather(c("NEP", "RE", "GPP"), key = "variable", value = "value") %>%  
+  filter(variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
+  mutate(description = paste0(variable, '\n (MgC ha-1 year-1)')) -> 
+  more_obs
+
 ED_output %>%  
@@
   geom_line(data = to_plot, aes(year, value, group = met), 
-            alpha = 0.9, color = "#999999", ) + 
+            alpha = 0.9, color = "#999999") + 
   geom_pointrange(data = obs, aes(x=year, y=mean, ymin=lower, ymax=upper)) + 
+  geom_line(data = more_obs, aes(Year, value), color = "black") + 
+  geom_point(data = chrono, aes(year, value), color = "red") + 
   facet_wrap("description", scales = "free", 
@@
   geom_point(alpha = 0.75, size = 1.5) + 
-  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
+  geom_smooth(method = "lm", size = 0.5, alpha = 0.2, show_guide  = FALSE) + 
   stat_cor(method = "pearson", 
@@
            label.y.npc = "bottom",
-           r.accuracy = 0.01) + 
+           r.accuracy = 0.01, 
+            show_guide  = FALSE) + 
   labs(x = "Resistance", 
@@
   plot
-  ggsave(plot, filename = "figures/resistance_resilience3.png", width = 6, height = 4.5)
+  ggsave(plot, 
+         filename = "figures/resistance_resilience3.png", width = 6, height = 4.5)
 
